### 2018-06-28 doc/plos/text/manuscript.tex
add title
It's a start, right? I used the title from my thesis, but we can still
pick a better one.

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2".
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Multiple Resolution Nowcasting of Influenza through Sensor Fusion} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Name1 Surname\textsuperscript{1,2\Yinyang},
Name2 Surname\textsuperscript{2\Yinyang},
Name3 Surname\textsuperscript{2,3\textcurrency},
Name4 Surname\textsuperscript{2},
Name5 Surname\textsuperscript{2\ddag},
Name6 Surname\textsuperscript{2\ddag},
Name7 Surname\textsuperscript{1,2,3*},
with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
%
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address
% \textcurrency c Insert third current address

% Deceased author note
\dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* correspondingauthor@institute.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\begin{eqnarray}
\label{eq:schemeP}
	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
\end{eqnarray}

\section*{Materials and methods}
\subsection*{Etiam eget sapien nibh}

% For figure citations, please use "Fig" instead of "Figure".
Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

% Results and Discussion can be combined.
\section*{Results}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place tables after the first paragraph in which they are cited.
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l+l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
\subsubsection*{3rd level heading}
Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat.

\begin{enumerate}
	\item{react}
	\item{diffuse free particles}
	\item{increment time by dt and go to 1}
\end{enumerate}

\subsection*{Sed ac quam id nisi malesuada congue}

Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{itemize}
	\item First bulleted item.
	\item Second bulleted item.
	\item Third bulleted item.
\end{itemize}

\section*{Discussion}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero~\cite{bib3}.

\section*{Conclusion}

CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 File.}
\label{S1_File}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Video.}
\label{S1_Video}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{S1_Appendix}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for
% step-by-step instructions.
%
\begin{thebibliography}{10}

\bibitem{bib1}
Conant GC, Wolfe KH.
\newblock {{T}urning a hobby into a job: how duplicated genes find new
  functions}.
\newblock Nat Rev Genet. 2008 Dec;9(12):938--950.

\bibitem{bib2}
Ohno S.
\newblock Evolution by gene duplication.
\newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
  Springer-Verlag.; 1970.

\bibitem{bib3}
Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
\newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
  infection through a transposon insertion followed by a {D}uplication}.
\newblock PLoS Genet. 2011 Oct;7(10):e1002337.

\end{thebibliography}

\end{document}

### 2018-06-28 doc/plos/text/manuscript.tex
use sentence case for title

@@
 {\Large
-\textbf\newline{Multiple Resolution Nowcasting of Influenza through Sensor Fusion} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Multiple resolution nowcasting of influenza through sensor fusion} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }

### 2018-06-28 doc/plos/text/manuscript.tex
remove sample content add references
Removed all of the sample stuff from the template, added a couple of
authors (feel free to add yourself if you're reading this), added
references.bib (currently just contains my thesis).

The PDF is here (for now):
https://delphi.midas.cs.cmu.edu/nowcast/plos/data/manuscript_20180628.pdf

@@
 \\
-Name1 Surname\textsuperscript{1,2\Yinyang},
-Name2 Surname\textsuperscript{2\Yinyang},
-Name3 Surname\textsuperscript{2,3\textcurrency},
-Name4 Surname\textsuperscript{2},
-Name5 Surname\textsuperscript{2\ddag},
-Name6 Surname\textsuperscript{2\ddag},
-Name7 Surname\textsuperscript{1,2,3*},
-with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
+David Farrow\textsuperscript{1},
+% Anyone else? Feel free to edit.
+Ryan Tibshirani\textsuperscript{2},
+Roni Rosenfeld\textsuperscript{1,2}
 \\
 \bigskip
-\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{1} Computational Biology Department, Carnegie Mellon University, Pittsburgh, PA, USA
 \\
-\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
-\\
-\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\textbf{2} Machine Learning Department, Carnegie Mellon University, Pittsburgh, PA, USA
 \\
@@
 
-% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
-%
-% Remove or comment out the author notes below if they aren't used.
-%
-% Primary Equal Contribution Note
-\Yinyang These authors contributed equally to this work.
-
-% Additional Equal Contribution Note
-% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
-\ddag These authors also contributed equally to this work.
-
-% Current address notes
-\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
-% \textcurrency b Insert second current address
-% \textcurrency c Insert third current address
-
-% Deceased author note
-\dag Deceased
-
-% Group/Consortium Author Note
-\textpilcrow Membership list can be found in the Acknowledgments section.
-
 % Use the asterisk to denote corresponding authorship and provide email address in note below.
-* correspondingauthor@institute.edu
+* roni.rosenfeld@cs.cmu.edu
 
@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+TODO
 
@@
 \section*{Author summary}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+TODO
 
@@
 \section*{Introduction}
-Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
-\begin{eqnarray}
-\label{eq:schemeP}
-	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
-\end{eqnarray}
+TODO~\cite{farrow2016modeling}.
 
 \section*{Materials and methods}
-\subsection*{Etiam eget sapien nibh}
 
-% For figure citations, please use "Fig" instead of "Figure".
-Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
-
-% Place figure captions after the first paragraph in which they are cited.
-\begin{figure}[!h]
-\caption{{\bf Bold the figure title.}
-Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
-\label{fig1}
-\end{figure}
+TODO
 
@@
 \section*{Results}
-Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
 
-% Place tables after the first paragraph in which they are cited.
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
-\centering
-\caption{
-{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
-\begin{tabular}{|l+l|l|l|l|l|l|l|}
-\hline
-\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
-$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
-$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
-$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
-\end{tabular}
-\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-\end{flushleft}
-\label{table1}
-\end{adjustwidth}
-\end{table}
-
-%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
-\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
-\subsubsection*{3rd level heading}
-Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
-Sed non aliquet felis.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris commodo justo ac dui pretium imperdiet.
-Sed suscipit iaculis mi at feugiat.
-
-\begin{enumerate}
-	\item{react}
-	\item{diffuse free particles}
-	\item{increment time by dt and go to 1}
-\end{enumerate}
-
-\subsection*{Sed ac quam id nisi malesuada congue}
-
-Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
-
-\begin{itemize}
-	\item First bulleted item.
-	\item Second bulleted item.
-	\item Third bulleted item.
-\end{itemize}
+TODO
 
 \section*{Discussion}
-Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero~\cite{bib3}.
+
+TODO
 
@@
 
-CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
-
-Sed non aliquet felis.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris commodo justo ac dui pretium imperdiet.
-Sed suscipit iaculis mi at feugiat.
-Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna.
-Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-Nam condimentum sem eget mollis euismod.
-Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex.
-Nunc est dui, sodales sed mauris nec, auctor sagittis leo.
-Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor.
-For more information, see \nameref{S1_Appendix}.
+TODO
 
@@
 
-% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
-\paragraph*{S1 Fig.}
-\label{S1_Fig}
-{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).
-
-\paragraph*{S2 Fig.}
-\label{S2_Fig}
-{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 File.}
-\label{S1_File}
-{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 Video.}
-\label{S1_Video}
-{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 Appendix.}
-\label{S1_Appendix}
-{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 Table.}
-\label{S1_Table}
-{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
+TODO
 
 \section*{Acknowledgments}
-Cras egestas velit mauris, eu mollis turpis pellentesque sit amet.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Nam id pretium nisi.
-Sed ac quam id nisi malesuada congue.
-Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.
+
+TODO
 
@@
 
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-% or
-%
 % Compile your BiBTeX database using our plos2015.bst
@@
 % step-by-step instructions.
-%
-\begin{thebibliography}{10}
 
-\bibitem{bib1}
-Conant GC, Wolfe KH.
-\newblock {{T}urning a hobby into a job: how duplicated genes find new
-  functions}.
-\newblock Nat Rev Genet. 2008 Dec;9(12):938--950.
+% Uncomment this to generate the *.bbl file.
+\bibliography{references}
 
-\bibitem{bib2}
-Ohno S.
-\newblock Evolution by gene duplication.
-\newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
-  Springer-Verlag.; 1970.
+% Paste the contents of the *.bbl file below this line.
+\begin{thebibliography}{1}
 
-\bibitem{bib3}
-Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
-\newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
-  infection through a transposon insertion followed by a {D}uplication}.
-\newblock PLoS Genet. 2011 Oct;7(10):e1002337.
+\bibitem{farrow2016modeling}
+Farrow DC.
+\newblock Modeling the Past, Present, and Future of Influenza [PhD Thesis].
+\newblock Carnegie Mellon University; 2016.
+\newblock Available from:
+  \url{http://reports-archive.adm.cs.cmu.edu/anon/cbd/CMU-CB-16-101.pdf}.
 
 \end{thebibliography}
+% Paste the contents of the *.bbl file above this line.
 

### 2018-06-29 doc/plos/text/manuscript.tex
add author summary

@@
 
-TODO
+Despite ongoing advances in public health, medicine, and related fields,
+influenza remains an ever-present burden to modern society. One way in which
+the impact of influenza outbreaks can be reduced is though greater situational
+awareness, leading to improvements in prevention and preparedness. The advent
+of digital surveillance and advances in the field of epidemiological
+forecasting have given rise to a diverse set of noisy proxies of influenza
+activity. It has been shown that a combination of several such proxies results
+in increased signal-to-noise ratio and overall more accurate nowcasts at the US
+national level. However, serious challenges remain in assimilating proxies
+which vary in geographic resolution and which may be intermittently available.
+To address these issues, we propose a sensor fusion approach to nowcasting
+influenza. In agreement with previous results, we find that aggregate nowcasts
+are overall more accurate than individual proxies. Furthermore, we produce
+nowcasts at the level of US states, even though most proxies are only available
+nationally or regionally. We make available ongoing, weekly nowcasts of
+influenza activity in the US---nationally, regionally, and for all states.
 

### 2018-06-29 doc/plos/text/manuscript.tex
mention supporting information
include author middle initials
add background with references

@@
 \\
-David Farrow\textsuperscript{1},
+David C. Farrow\textsuperscript{1},
 % Anyone else? Feel free to edit.
-Ryan Tibshirani\textsuperscript{2},
+Ryan J. Tibshirani\textsuperscript{2},
 Roni Rosenfeld\textsuperscript{1,2}
@@
 
-TODO~\cite{farrow2016modeling}.
+\subsection*{Background}
+
+Annual influenza epidemics incur tremendous expenses and are strongly
+detrimental to societal wellbeing \cite{szucs1999socio, Stohr2002,
+molinari2007annual, who_flu_impact}. One way in which these impacts can be
+lessened is though outbreak forewarning and therefore improved preparedness
+\cite{myers2000}. This has been a primary impetus in the development of a large
+number of influenza forecasting frameworks in recent years \cite{nsoesie2013a,
+chretien2014, brooks2015, farrow2017human}. TODO(cite new brooks, cdc papers)
+
+Forecasting critically depends on timely and accurate situational awareness,
+and in the United States (US) one widespread measure of influenza burden is the
+population-weighted percent influenza-like illness (wILI), derived from the US
+Outpatient Influenza-Like Illness Surveillance Network (ILINet). On a weekly
+basis, health care providers who have volunteered to participate in the ILINet
+program submit, among other data, the percent of their cases that have met the
+syndromic criteria of Influenza-Like Illness (\%ILI)---fever with either cough
+or sore throat, without a known cause other than influenza \cite{wilidef,
+who_ili_def}. The US Centers for Disease Control and Prevention (CDC) publishes
+weekly wILI---a weighted combination of state \%ILI---for several locations
+within the US, including the ten Health and Human Services (HHS) regions, the
+nine census divisions, and the majority of US states and territories.
+
+Although \%ILI, and by extension wILI, is highly beneficial in a number of
+settings \cite{ritzwoller2005comparison, thompson2006epidemiology,
+m2008infectious}, it is ill-suited for real-time estimation of influenza
+activity for several reasons. First, there is an inherent delay of one or more
+weeks between patient care and wILI availability. In the typical and best case,
+data collected between Sunday and Saturday of some week is reported on Friday
+of the following week. Second, wILI is subject to retrospective and significant
+revision due to a process we refer to as ``backfill''. Backfill is the eventual
+incorporation of reports submitted multiple weeks after patient care, and with
+the additional data comes a different, but more representative, value of wILI.
+Third, wILI is aggregated across broad geographic regions before publication.
+It is well known, however, that the spread of influenza---particularly around
+the peak of the epidemic---is spatially heterogeneous between and within states
+\cite{bonabeau1998geographical, viboud2006synchrony, eggo2010spatial,
+viboud2013contrasting, gog2014spatial}, and therefore regional, divisional, and
+especially national wILI is a convolution of influenza activity within
+constituent states.
+
+TODO
 
@@
 
-TODO
+TODO~\cite{farrow2016modeling}
 
@@
 
-TODO
+\paragraph*{S1 Text}
+\label{S1_Text}
+{\bf Supporting Information.} Includes derivation and sensitivity analysis.
 
@@
 % Uncomment this to generate the *.bbl file.
-\bibliography{references}
+%\bibliography{references}
 
 % Paste the contents of the *.bbl file below this line.
-\begin{thebibliography}{1}
+\begin{thebibliography}{10}
+
+\bibitem{szucs1999socio}
+Szucs T.
+\newblock The socio-economic burden of influenza.
+\newblock Journal of Antimicrobial Chemotherapy. 1999;44(suppl 2):11--15.
+
+\bibitem{Stohr2002}
+St{\"o}hr K.
+\newblock Influenza--WHO cares.
+\newblock The Lancet Infectious diseases. 2002;2(9):517--517.
+
+\bibitem{molinari2007annual}
+Molinari NAM, Ortega-Sanchez IR, Messonnier ML, Thompson WW, Wortley PM,
+  Weintraub E, et~al.
+\newblock The annual impact of seasonal influenza in the US: measuring disease
+  burden and costs.
+\newblock Vaccine. 2007;25(27):5086--5096.
+
+\bibitem{who_flu_impact}
+Influenza (Seasonal); 2018.
+\newblock Available from:
+  \url{http://www.who.int/mediacentre/factsheets/fs211/en/}.
+
+\bibitem{myers2000}
+Myers MF, Rogers D, Cox J, Flahault A, Hay S.
+\newblock Forecasting disease risk for increased epidemic preparedness in
+  public health.
+\newblock Advances in Parasitology. 2000;47:309--330.
+
+\bibitem{nsoesie2013a}
+Nsoesie EO, Brownstein JS, Ramakrishnan N, Marathe MV.
+\newblock A systematic review of studies on forecasting the dynamics of
+  influenza outbreaks.
+\newblock Influenza Other Respi Viruses. 2013;8(3):309--316.
+\newblock doi:{10.1111/irv.12226}.
+
+\bibitem{chretien2014}
+Chretien JP, George D, Shaman J, Chitale RA, McKenzie FE.
+\newblock Influenza Forecasting in Human Populations: A Scoping Review.
+\newblock {PLoS} {ONE}. 2014;9(4):e94130.
+\newblock doi:{10.1371/journal.pone.0094130}.
+
+\bibitem{brooks2015}
+Brooks LC, Farrow DC, Hyun S, Tibshirani RJ, Rosenfeld R.
+\newblock Flexible Modeling of Epidemics with an Empirical Bayes Framework.
+\newblock {PLOS} Computational Biology. 2015;11(8):e1004382.
+\newblock doi:{10.1371/journal.pcbi.1004382}.
+
+\bibitem{farrow2017human}
+Farrow DC, Brooks LC, Hyun S, Tibshirani RJ, Burke DS, Rosenfeld R.
+\newblock A human judgment approach to epidemiological forecasting.
+\newblock PLOS Computational Biology. 2017;13(3):e1005248.
+
+\bibitem{wilidef}
+Overview of Influenza Surveillance in the United States; 2015.
+\newblock Available from: \url{http://www.cdc.gov/flu/weekly/overview.htm}.
+
+\bibitem{who_ili_def}
+WHO surveillance case definitions for ILI and SARI; 2014.
+\newblock Available from:
+  \url{http://www.who.int/influenza/surveillance_monitoring/ili_sari_surveillance_case_definition/en/}.
+
+\bibitem{ritzwoller2005comparison}
+Ritzwoller DP, Kleinman K, Palen T, Abrams A, Kaferly J, Yih W, et~al.
+\newblock Comparison of syndromic surveillance and a sentinel provider system
+  in detecting an influenza outbreak—Denver, Colorado, 2003.
+\newblock MMWR Morb Mortal Wkly Rep. 2005;54(Suppl.):151--6.
+
+\bibitem{thompson2006epidemiology}
+Thompson WW, Comanor L, Shay DK.
+\newblock Epidemiology of seasonal influenza: use of surveillance data and
+  statistical models to estimate the burden of disease.
+\newblock Journal of Infectious Diseases. 2006;194(Supplement 2):S82--S91.
+
+\bibitem{m2008infectious}
+M'ikanatha NM, Lynfield R, Van~Beneden CA, de~Valk H.
+\newblock Infectious disease surveillance.
+\newblock John Wiley \& Sons; 2008.
+
+\bibitem{bonabeau1998geographical}
+Bonabeau E, Toubiana L, Flahault A.
+\newblock The geographical spread of influenza.
+\newblock Proceedings of the Royal Society of London B: Biological Sciences.
+  1998;265(1413):2421--2425.
+
+\bibitem{viboud2006synchrony}
+Viboud C, Bj{\o}rnstad ON, Smith DL, Simonsen L, Miller MA, Grenfell BT.
+\newblock Synchrony, waves, and spatial hierarchies in the spread of influenza.
+\newblock science. 2006;312(5772):447--451.
+
+\bibitem{eggo2010spatial}
+Eggo RM, Cauchemez S, Ferguson NM.
+\newblock Spatial dynamics of the 1918 influenza pandemic in England, Wales and
+  the United States.
+\newblock Journal of The Royal Society Interface. 2010; p. rsif20100216.
+
+\bibitem{viboud2013contrasting}
+Viboud C, Nelson MI, Tan Y, Holmes EC.
+\newblock Contrasting the epidemiological and evolutionary dynamics of
+  influenza spatial transmission.
+\newblock Philosophical Transactions of the Royal Society of London B:
+  Biological Sciences. 2013;368(1614):20120199.
+
+\bibitem{gog2014spatial}
+Gog JR, Ballesteros S, Viboud C, Simonsen L, Bjornstad ON, Shaman J, et~al.
+\newblock Spatial transmission of 2009 pandemic influenza in the US.
+\newblock PLoS Comput Biol. 2014;10(6):e1003635.
 

### 2018-06-29 doc/plos/text/manuscript.tex
add related work subsection

@@
 
+\subsection*{Related Work}
+
+An approach often used to address some of these issues is to incorporate novel
+data streams (NDS) \cite{althouse2015enhancing}. These signals, when based on
+search queries \cite{eysenbach2006infodemiology, polgreen2008using,
+hulth2009web, ginsberg2009, dugas2013, araz2014using, santillana2014can,
+preis2014adaptive}, social media \cite{ritterman2009, culotta2010towards,
+signorini2011use, achrekar2011predicting, broniatowski2013national,
+paul2014twitter}, information-seeking behavior \cite{generous2014global,
+mciver2014wikipedia, hickmann2015forecasting}, and participatory surveillance
+\cite{smolinski2015flu, chunara2013flu}, are collectively termed ``digital
+surveillance'', as opposed to traditional surveillance of which wILI is one
+example. Although many of these signals are available with little to no lag,
+each has distinct shortcomings, and it seems that common practice has been to
+nowcast wILI using any one of the available digital surveillance signals in
+isolation. However, it was more recently demonstrated that multiple digital
+surveillance signals can be used in parallel to produce more accurate nowcasts
+of national wILI \cite{santillana2015combining}.
+
 TODO
@@
 
+\bibitem{althouse2015enhancing}
+Althouse BM, Scarpino SV, Meyers LA, Ayers JW, Bargsten M, Baumbach J, et~al.
+\newblock Enhancing disease surveillance with novel data streams: challenges
+  and opportunities.
+\newblock EPJ Data Science. 2015;4(1):1--8.
+
+\bibitem{eysenbach2006infodemiology}
+Eysenbach G.
+\newblock Infodemiology: tracking flu-related searches on the web for syndromic
+  surveillance.
+\newblock In: AMIA Annual Symposium Proceedings. vol. 2006. American Medical
+  Informatics Association; 2006. p. 244.
+
+\bibitem{polgreen2008using}
+Polgreen PM, Chen Y, Pennock DM, Nelson FD, Weinstein RA.
+\newblock Using internet searches for influenza surveillance.
+\newblock Clinical infectious diseases. 2008;47(11):1443--1448.
+
+\bibitem{hulth2009web}
+Hulth A, Rydevik G, Linde A.
+\newblock Web queries as a source for syndromic surveillance.
+\newblock PloS one. 2009;4(2):e4378.
+
+\bibitem{ginsberg2009}
+Ginsberg J, Mohebbi MH, Patel RS, Brammer L, Smolinski MS, Brilliant L.
+\newblock Detecting influenza epidemics using search engine query data.
+\newblock Nature. 2009;457(7232):1012--1014.
+
+\bibitem{dugas2013}
+Dugas AF, Jalalpour M, Gel Y, Levin S, Torcaso F, Igusa T, et~al.
+\newblock Influenza Forecasting with Google Flu Trends.
+\newblock {PLoS} {ONE}. 2013;8(2):e56176.
+\newblock doi:{10.1371/journal.pone.0056176}.
+
+\bibitem{araz2014using}
+Araz OM, Bentley D, Muelleman RL.
+\newblock Using Google Flu Trends data in forecasting influenza-like--illness
+  related ED visits in Omaha, Nebraska.
+\newblock The American journal of emergency medicine. 2014;32(9):1016--1023.
+
+\bibitem{santillana2014can}
+Santillana M, Zhang DW, Althouse BM, Ayers JW.
+\newblock What can digital disease detection learn from (an external revision
+  to) Google Flu Trends?
+\newblock American journal of preventive medicine. 2014;47(3):341--347.
+
+\bibitem{preis2014adaptive}
+Preis T, Moat HS.
+\newblock Adaptive nowcasting of influenza outbreaks using Google searches.
+\newblock Royal Society open science. 2014;1(2):140095.
+
+\bibitem{ritterman2009}
+Ritterman J, Osborne M, Klein E.
+\newblock Using prediction markets and Twitter to predict a swine flu pandemic.
+\newblock In: 1st international workshop on mining social media. vol.~9. ac.
+  uk/miles/papers/swine09. pdf (accessed 26 August 2015); 2009. p. 9--17.
+
+\bibitem{culotta2010towards}
+Culotta A.
+\newblock Towards detecting influenza epidemics by analyzing Twitter messages.
+\newblock In: Proceedings of the first workshop on social media analytics. ACM;
+  2010. p. 115--122.
+
+\bibitem{signorini2011use}
+Signorini A, Segre AM, Polgreen PM.
+\newblock The use of Twitter to track levels of disease activity and public
+  concern in the US during the influenza A H1N1 pandemic.
+\newblock PloS one. 2011;6(5):e19467.
+
+\bibitem{achrekar2011predicting}
+Achrekar H, Gandhe A, Lazarus R, Yu SH, Liu B.
+\newblock Predicting flu trends using twitter data.
+\newblock In: Computer Communications Workshops (INFOCOM WKSHPS), 2011 IEEE
+  Conference on. IEEE; 2011. p. 702--707.
+
+\bibitem{broniatowski2013national}
+Broniatowski DA, Paul MJ, Dredze M.
+\newblock National and local influenza surveillance through Twitter: an
+  analysis of the 2012-2013 influenza epidemic.
+\newblock PloS one. 2013;8(12):e83672.
+
+\bibitem{paul2014twitter}
+Paul MJ, Dredze M, Broniatowski D.
+\newblock Twitter improves influenza forecasting.
+\newblock PLoS currents. 2014;6.
+
+\bibitem{generous2014global}
+Generous N, Fairchild G, Deshpande A, Del~Valle SY, Priedhorsky R.
+\newblock Global disease monitoring and forecasting with Wikipedia.
+\newblock PLoS Comput Biol. 2014;10(11):e1003892.
+
+\bibitem{mciver2014wikipedia}
+McIver DJ, Brownstein JS.
+\newblock Wikipedia usage estimates prevalence of influenza-like illness in the
+  United States in near real-time.
+\newblock PLoS Comput Biol. 2014;10(4):e1003581.
+
+\bibitem{hickmann2015forecasting}
+Hickmann KS, Fairchild G, Priedhorsky R, Generous N, Hyman JM, Deshpande A,
+  et~al.
+\newblock Forecasting the 2013--2014 influenza season using Wikipedia.
+\newblock PLoS Comput Biol. 2015;11(5):e1004239.
+
+\bibitem{smolinski2015flu}
+Smolinski MS, Crawley AW, Baltrusaitis K, Chunara R, Olsen JM, W{\'o}jcik O,
+  et~al.
+\newblock Flu Near You: Crowdsourced Symptom Reporting Spanning 2 Influenza
+  Seasons.
+\newblock American journal of public health. 2015;105(10):2124--2130.
+
+\bibitem{chunara2013flu}
+Chunara R, Aman S, Smolinski M, Brownstein JS.
+\newblock Flu near you: an online self-reported influenza surveillance system
+  in the USA.
+\newblock Online Journal of Public Health Informatics. 2013;5(1).
+
+\bibitem{santillana2015combining}
+Santillana M, Nguyen AT, Dredze M, Paul MJ, Nsoesie EO, Brownstein JS.
+\newblock Combining search, social media, and traditional data sources to
+  improve influenza surveillance.
+\newblock PLoS Comput Biol. 2015;11(10):e1004513.
+
 \bibitem{farrow2016modeling}

### 2018-06-29 doc/plos/text/manuscript.tex
add contribution subsection

@@
 
-TODO
+\subsection*{Contribution}
+
+We propose nowcasting influenza through \emph{sensor
+fusion}~\cite{farrow2016modeling} of digital surveillance. Derived from the
+Kalman filter \cite{kalman1960new}, sensor fusion is optimal in a least-squares
+sense (up to assumptions of independence and normality), robust to noise,
+tolerant of missing data, and capable of assimilating data across varying
+geospatial resolutions. We implement a sensor fusion framework and apply it to
+nowcasting ILI in US states and territories (and in any combination thereof,
+including HHS regions, census divisions, and the US as a whole), based on the
+assimilation of any and all surveillance data available at runtime.
 
@@
 
-TODO~\cite{farrow2016modeling}
+TODO
 
@@
 
+\bibitem{kalman1960new}
+Kalman RE.
+\newblock A new approach to linear filtering and prediction problems.
+\newblock Journal of basic Engineering. 1960;82(1):35--45.
+
 \end{thebibliography}

### 2018-07-02 doc/plos/text/manuscript.tex
add subsection for data sources

@@
 
+\subsection*{Data Sources}
+
+Data sources used in nowcasting can be broadly separated into two classes:
+ground truth and surveillance. The aim at any given point in time is to predict
+ground truth, given surveillance. Surveillance can be further divided into
+digital surveillance (e.g. based on social media) and predictive surveillance
+(e.g. based on short-term forecasts). Below are the specific data sources we
+use in each category.
+
+\subsubsection*{Ground Truth}
+
+\begin{description}
+  \item{\textbf{Population weighted percent Influenza-like Illness (wILI)}} \\
+    Each week, CDC publishes wILI for the US as a whole, the 10 HHS regions,
+    the 9 census divisions, and most states and territories \cite{wilisrc}. In
+    any given location, wILI is the mean (weighted by population) of ILI in
+    constituent locations. ILI is the percent of influenza-like illness out of
+    all patient visits voluntarily reported by participating healthcare
+    providers. Because reports continue to be integrated after initial
+    publication, wILI is subject to retrospective revision in a process we
+    refer to as ``backfill''. As a result, wILI is effectively versioned. When
+    training models and making predictions, we use preliminary---not
+    final---values of wILI whenever possible.
+\end{description}
+
+\subsubsection*{Digital Surveillance}
+
+\begin{description}
+  \item{\textbf{Google Flu Trends}} \\
+    Google Flu Trends was a weekly prediction of wILI based on Google searches.
+    The service was originally released in 2008; was updated in 2009, 2013, and
+    2014 \cite{ginsberg2009, gft_main}; and was discontinued in 2015.
+  \item{\textbf{Google Health Trends}} \\
+    Google Health Trends is an API by Google for research purposes which makes
+    available, among many other things, daily counts (on an arbitrary scale)
+    for a predefined influenza topic. TODO(link)
+  \item{\textbf{HealthTweets}} \\
+    HealthTweets (http://healthtweets.org) makes available the daily volume of
+    tweets indicative of ILI \cite{broniatowski2013national,
+    dredze2014healthtweets}.
+  \item{\textbf{Wikimedia Foundation}} \\
+    The Wikimedia Foundation makes available the number of hits, per hour, to
+    every Wikipedia article. TODO(link)
+  \item{\textbf{CDC Page Hits}} \\
+    CDC makes available the number of daily hits per page, per state for their
+    website (https:/www.cdc.gov/).
+\end{description}
+
+\subsubsection*{Predictive Surveillance}
+
+\begin{description}
+  \item{\textbf{Epicast (EPIC)}} \\
+    Epicast \cite{farrow2017human} is an influenza forecasting system based on
+    collective human judgment. From these forecasts, we interpret the
+    one-week-ahead value of wILI as a nowcast. TODO(link)
+  \item{\textbf{Seasonal Autoregression (SAR3)}} \\
+    SAR3 \cite{farrow2016modeling} provides an autoregression estimate of wILI
+    on the current week. TODO(link)
+  \item{\textbf{Epidemic Archetype (ARCH)}} \\
+    ARCH \cite{farrow2016modeling} is an empirical model based on the Empirical
+    Bayes forecasting framework \cite{brooks2015}. Like Epicast, ARCH provides
+    a short-term epidemiological forecast from which we interpret the
+    one-week-ahead prediction as a nowcast.
+\end{description}
+
+With the exception of Google Health Trends, these data sources are publicly
+available via the Delphi Epidata API. TODO(can we share TWTR? GHT?) TODO(link)
+
+\subsection*{Fitting}
+
+%As others have done for flu and other infectious diseases
+%\cite{generous2014global, hickmann2015forecasting, mciver2014wikipedia}, we use
+%the popularity of selected articles to estimate wILI for the US as a whole.
+%
+%As with WIKI, we use popularity of selected pages to estimate ILI within US
+%states.
+%
+%Inputs available at sub-weekly temporal resolution are aggregated to form
+%weekly readings.
+
 TODO
@@
 
+\bibitem{wilisrc}
+Influenza National and Regional Level Graphs and Data; 2017.
+\newblock Available from:
+  \url{https://gis.cdc.gov/grasp/fluview/fluportaldashboard.html}.
+
+\bibitem{gft_main}
+Google Flu Trends; 2015.
+\newblock Available from: \url{https://www.google.org/flutrends/about/}.
+
+\bibitem{dredze2014healthtweets}
+Dredze M, Cheng R, Paul MJ, Broniatowski D.
+\newblock HealthTweets.org: a platform for public health surveillance using
+  twitter.
+\newblock In: AAAI Conference on Artificial Intelligence; 2014. p. 1--2.
+
 \end{thebibliography}

### 2018-07-02 doc/plos/code/make_data_index.py
creation date is mangled, don't show it

@@ -14,7 +14,7 @@ def main():
     if os.path.isdir(item):
       item += '/'
     link = '<a href="%s">%s</a>' % (item, item)
-    print('<li>%s<br>%.2f MiB; Created %s</li>' % (link, size, when))
+    print('<li>%s<br>%.2f MiB</li>' % (link, size))
   print('</ul></html>')
 
 

### 2018-07-03 doc/plos/text/manuscript.tex
add sensor fitting subsection

@@
 \begin{description}
-  \item{\textbf{Google Flu Trends}} \\
+  \item{\textbf{Google Flu Trends (GFT)}} \\
     Google Flu Trends was a weekly prediction of wILI based on Google searches.
@@
     2014 \cite{ginsberg2009, gft_main}; and was discontinued in 2015.
-  \item{\textbf{Google Health Trends}} \\
+  \item{\textbf{Google Health Trends (GHT)}} \\
     Google Health Trends is an API by Google for research purposes which makes
@@
     for a predefined influenza topic. TODO(link)
-  \item{\textbf{HealthTweets}} \\
+  \item{\textbf{HealthTweets (TWTR)}} \\
     HealthTweets (http://healthtweets.org) makes available the daily volume of
@@
     dredze2014healthtweets}.
-  \item{\textbf{Wikimedia Foundation}} \\
+  \item{\textbf{Wikimedia Foundation (WIKI)}} \\
     The Wikimedia Foundation makes available the number of hits, per hour, to
     every Wikipedia article. TODO(link)
-  \item{\textbf{CDC Page Hits}} \\
+  \item{\textbf{CDC Page Hits (CDCP)}} \\
     CDC makes available the number of daily hits per page, per state for their
@@
 
-%As others have done for flu and other infectious diseases
-%\cite{generous2014global, hickmann2015forecasting, mciver2014wikipedia}, we use
-%the popularity of selected articles to estimate wILI for the US as a whole.
-%
-%As with WIKI, we use popularity of selected pages to estimate ILI within US
-%states.
-%
-%Inputs available at sub-weekly temporal resolution are aggregated to form
-%weekly readings.
+To obtain a signal usable for nowcasting, we must fit inputs to wILI. Since
+this is trivially the case for predictive surveillance (i.e. EPIC, SAR3, and
+ARCH), we use those values directly. For digital survaillance, we use
+historical data to build a linear regression model mapping raw inputs to wILI.
+We then apply this model with the most recent data to obtain a prediction of
+wILI. We repeat the fitting procedure each week prior to nowcasting to obtain a
+new model based on the most up-to-date data. The fitting procedure is described
+in more detail in \cite{farrow2016modeling}.
 
-TODO
+Data sources with a single covariate (i.e. GFT, GHT, TWTR) are fit with
+ordinary linear regression. Data sources with multiple covariates (i.e. WIKI,
+CDCP) are fit with multiple linear regression.
+
+As others have done for flu and other infectious diseases
+\cite{generous2014global, hickmann2015forecasting, mciver2014wikipedia}, we use
+the volume of Wikipedia traffic to estimate wILI for the US as a whole. To
+produce a signal, we fit seven covariates to wILI, each corresponding to a
+particular flu-releated Wikipedia article. The articles TODO(list in SI) were
+chosen offline and by hand \textit{a priori}. Since WIKI has no attached
+location information, we filter by particular hours of the day TODO(list in SI)
+as a very rough proxy of Wikipedia activity in the US. Although this is a
+grossly invalid assumption, preliminary investigation showed this to provide a
+better fit to wILI.
+
+Similarly, CDCP consists of six covariates, each corresponding to a particular
+flu-releated page on the CDC website. The pages TODO(list in SI) were chosen
+offline and by hand \textit{a priori}. CDCP is unique in that the covariates
+are log-transformed before fitting as preliminary investigation showed this to
+provide a better fit to wILI.
+
+For all digital surveillance sources, we apply the fitting procedure separately
+for each available location (e.g. US as a whole, HHS regions, census divisions,
+and US states).
+
+For data sources with sub-weekly temporal resolution (i.e. GHT, TWTR, WIKI,
+CDCP), we aggregate counts by week before applying the fitting procedure. In
+future work we aim to incorporate these sources as soon as they are available,
+rather than waiting until the end of the week to use their aggregate values.
 

### 2018-07-03 doc/plos/text/manuscript.tex
add fusion subsection

@@
 
+\subsection*{Fusion}
+
+The fitting procedure described above gives a disparate set of predictions of
+wILI for various locations in the US. We combine---via sensor fusion
+\cite{farrow2016modeling}---all available predictions into a single, unified
+nowcast of wILI throughout the US.
+
+Sensor fusion is a generalization of the Kalman filter \cite{kalman1960new}
+which relaxes assumptions regarding the process being tracked---the
+distribution of wILI in this case. In particular, sensor fusion describes the
+combination of a set of measurements without requiring a linear Markov process
+model. This combination is optimal in a least-squares sense when the
+measurements are unbiased with normally distributed error.
+
+For nowcasting flu, this means practically that the fusion inputs should be in
+the same units and on the same scale as wILI and should have normally
+distributed error. By applying the fitting procedure above, we meet the first
+requirement, and the assumption that errors are normally distributed is
+generally not grossly violated.
+
+Nowcasting is conceptually a two-step process. First, measurements are fused to
+give an estimate of latent system state (namely, wILI in US states and
+territories). Second, nowcasts for each location of interest (namely, wILI in
+the US as a whole, HHS regions, census divisions, states, and territories) are
+extracted as a linear combination of latent state wILI.
+
+More concretely, given a vector of ``sensor`` readings (surveillance values as
+described above) $z$, we apply the sensor fusion kernel to produce a vector
+nowcast $y$.
+
+\begin{align*}
+A &= (H^T R^{-1} H)^{-1} H^T R^{-1} \\
+x &= A z \\
+y &= W x
+\end{align*}
+
+Elements of $z$ are the sensor reading of a particular data source in a
+particular location, and elements of $y$ are a unified nowcast in each location
+of interest.
+
+Matrices $H$ and $W$ are mappings from latent state-space to measurement-space
+and output-space respectively. Matrix $R$ is the empirical covariance of sensor
+noise, where noise is defined as the difference between sensor readings and
+ground truth measured on historical data. Intermediate matrix $A$ conceptually
+represents the optimal translation from measurement-space to latent
+state-space, given the reverse mapping ($H$) and sensor noise covariance ($R$).
+
+$H$ and $W$ are defined by the geographical makeup and population distribution
+within the US; they are constants, modulo changes in population over long
+periods of time (e.g. years). On any given week, sensors $z$ are given by
+surveillance data, and covariance $R$ is estimated based on historical data.
+
 % Results and Discussion can be combined.

### 2018-07-06 doc/plos/text/manuscript.tex
add covariance subsection

@@
 
+\subsection*{Covariance}
+
+Given that multiple surveillance sources provide data at the level of US
+states, the total number of source-location pairs (``sensors'') is well into
+the hundreds. Their resulting covariance is very high-dimensional, especially
+relative to the amount of weekly historical training data that is typically
+available (e.g. a couple of years). Further, some data sources are only
+available seasonally (i.e. EPIC, ARCH), and some are relatively lacking in
+training data (i.e. CDCP, EPIC). In general, partial and missing data is a
+possibility for any data source.
+
+Estimating covariance in this situation is challenging, and while the
+statistical literature suggests several approaches TODO(cite a few of them),
+consensus appears to be lacking.
+
+We begin by computing the empirical pair-wise covariance of all sensors.
+However, the resulting matrix does not necessarily describe multi-dimensional
+covariance (i.e. it may not be positive definite). To correct this, we shrink
+this matrix toward the diagonal matrix of sensor variance. The magnitude of the
+shrinkage weight is chosen to maximize the multivariate normal log-likelihood
+of the available data. We subsequently use the shrunken matrix as $R$ in sensor
+fusion.
+
 % Results and Discussion can be combined.

### 2018-07-06 doc/plos/code/download_data.py
add script to download all data

@@ -1,0 +1,54 @@
+"""
+This is a one-off script to download ground truth (fluview) and sensors from
+the delphi api. Nowcasts are generated by nowcast_experiment.py. It's best to
+run this on the server since there is considerable latency when running this
+remotely.
+
+python3 -m nowcast.doc.plos.code.download_data
+"""
+
+
+from delphi.epidata.client.delphi_epidata import Epidata
+from delphi.nowcast.util.flu_data_source import FluDataSource
+from delphi.operations import secrets
+from delphi.utils.geo.locations import Locations
+
+
+weeks = Epidata.range(201030, 201824)
+
+
+def download_fluview(f):
+  for loc in Locations.region_list:
+    print('fluview', loc)
+    resp = Epidata.fluview(loc, weeks, auth=secrets.api.fluview)
+    rows = Epidata.check(resp)
+    for row in rows:
+      week, value = row['epiweek'], row['wili']
+      f.write('%d,%s,%.5f\n' % (week, loc, value))
+
+
+def download_sensors(f):
+  for sensor in FluDataSource.SENSORS:
+    for loc in Locations.region_list:
+      print(sensor, loc)
+      resp = Epidata.sensors(secrets.api.sensors, sensor, loc, weeks)
+      if resp['result'] not in (1, -2):
+        raise Exception(resp['result'])
+      if resp['result'] == -2:
+        print('(no data)')
+        continue
+      for row in resp['epidata']:
+        week, value = row['epiweek'], row['value']
+        f.write('%d,%s,%s,%.5f\n' % (week, sensor, loc, value))
+
+
+def main():
+  print('downloading... this may take a while (1m1.219s to be exact)')
+  with open('fluview.csv', 'w') as f:
+    download_fluview(f)
+  with open('sensors.csv', 'w') as f:
+    download_sensors(f)
+
+
+if __name__ == '__main__':
+  main()

### 2018-07-09 doc/plos/code/download_data.py
download unstable wili

@@ -27,6 +27,16 @@ def download_fluview(f):
       f.write('%d,%s,%.5f\n' % (week, loc, value))
 
 
+def download_preliminary_fluview(f):
+  for lag in range(3):
+    print('preliminary fluview', lag)
+    resp = Epidata.fluview('nat', weeks, lag=lag, auth=secrets.api.fluview)
+    rows = Epidata.check(resp)
+    for row in rows:
+      week, value = row['epiweek'], row['wili']
+      f.write('%d,%s,%.5f\n' % (week, 'nat_%d' % lag, value))
+
+
 def download_sensors(f):
   for sensor in FluDataSource.SENSORS:
     for loc in Locations.region_list:
@@ -46,6 +56,8 @@ def main():
   print('downloading... this may take a while (1m1.219s to be exact)')
   with open('fluview.csv', 'w') as f:
     download_fluview(f)
+  with open('fluview_prelim.csv', 'w') as f:
+    download_preliminary_fluview(f)
   with open('sensors.csv', 'w') as f:
     download_sensors(f)
 

### 2018-07-09 doc/plos/code/analyze.py
rudimentary plotting
Includes the sensor heatmap, which will probably be figure 1.

This is all quick and dirty so I can get a preview out faster. Pretty
plotting will be implemented later.

@@ -1,1 +1,107 @@
-# TODO: analyze the data and generate figures
+import csv
+import glob
+import os
+
+import numpy as np
+import matplotlib.pylab as plt
+
+from delphi.nowcast.util.flu_data_source import FluDataSource
+from delphi.utils.geo.locations import Locations
+from delphi.utils import epiweek as Epiweek
+
+from ugly_plot import UglyPlot
+
+
+class Analysis:
+
+  @staticmethod
+  def load_everything():
+    data = {}
+    for name in glob.glob('data/*.csv'):
+      key = os.path.split(name)[1][:-4]
+      data[key] = []
+      with open(name) as f:
+        reader = csv.reader(f)
+        for row in reader:
+          row[0] = int(row[0])
+          row[-1] = float(row[-1])
+          try:
+            row[-2] = float(row[-2])
+          except ValueError:
+            pass
+          data[key].append(row)
+    return data
+
+  def __init__(self):
+    self.data = Analysis.load_everything()
+
+  def get_truth(self, loc):
+    result = {}
+    for week, place, value in self.data['fluview']:
+      if place == loc:
+        result[week] = value
+    return result
+
+  def get_preliminary_truth(self, lag):
+    result = {}
+    for week, place, value in self.data['fluview_prelim']:
+      if place == 'nat_%d' % lag:
+        result[week] = value
+    return result
+
+  def get_sensor(self, name, loc):
+    result = {}
+    for week, sensor, place, value in self.data['sensors']:
+      if sensor == name and place == loc:
+        result[week] = value
+    return result
+
+  def get_experiment(self, name, loc):
+    result = {}
+    for week, place, value, std in self.data['nc_%s' % name]:
+      if place == loc:
+        result[week] = value
+    return result
+
+  def get_nowcast(self, loc):
+    return self.get_experiment('vanilla', loc)
+
+  def merge(self, a, b):
+    return dict((ew, (a[ew], b[ew])) for ew in a.keys() & b.keys())
+
+  def get_stats(self, merged):
+    a = np.array([merged[ew][0] for ew in merged])
+    b = np.array([merged[ew][1] for ew in merged])
+    return {
+      'n': len(merged),
+      'rmse': np.sqrt(np.mean(np.square(a - b))),
+      'mae': np.mean(np.abs(a - b)),
[…]

### 2018-07-09 doc/plos/code/ugly_plot.py

@@ -1,0 +1,21 @@
+# quick and dirty plots for development
+
+import numpy as np
+import matplotlib.pylab as plt
+
+
+class UglyPlot:
+
+  def __init__(self, analysis):
+    self.analysis = analysis
+
+  def plot_sensor_heatmap(self):
+    data, sensors, weeks = self.analysis.get_heatmap_data()
+    fig, ax = plt.subplots()
+    wk_idx = list(range(0, len(weeks), 100))
+    plt.xticks(wk_idx)
+    plt.yticks([0.5 + i for i in range(len(sensors))])
+    ax.set_xticklabels(['%d' % weeks[i] for i in wk_idx])
+    ax.set_yticklabels(sensors[::-1])
+    plt.pcolor(data[::-1, :])
+    plt.savefig('sensor_heatmap.png', bbox_inches='tight')

### 2018-07-09 doc/plos/code/analyze.py
sensor timeseries figure
This is the bottom panel of the first figure.

@@ -100,7 +100,7 @@ class Analysis:
 def main():
   analysis = Analysis()
   plotter = UglyPlot(analysis)
-  ploter.plot_sensor_heatmap()
+  plotter.plot_sensor_heatmap()
 
 
 if __name__ == '__main__':

### 2018-07-09 doc/plos/code/ugly_plot.py

@@ -9,6 +9,10 @@ class UglyPlot:
   def __init__(self, analysis):
     self.analysis = analysis
 
+  def _save(self, name):
+    plt.savefig('data/figures/' + name + '.png', bbox_inches='tight')
+    plt.close()
+
   def plot_sensor_heatmap(self):
     data, sensors, weeks = self.analysis.get_heatmap_data()
     fig, ax = plt.subplots()
@@ -18,4 +22,12 @@ class UglyPlot:
     ax.set_xticklabels(['%d' % weeks[i] for i in wk_idx])
     ax.set_yticklabels(sensors[::-1])
     plt.pcolor(data[::-1, :])
-    plt.savefig('sensor_heatmap.png', bbox_inches='tight')
+    self._save('sensor_heatmap_top')
+
+    for i, sensor in enumerate(sensors):
+      plt.plot(data[i, :], label=sensor)
+    t = self.analysis.get_truth('nat')
+    plt.plot([t[w] for w in weeks], label='Ground Truth', color='black')
+    plt.ylim([0, 10])
+    plt.legend()
+    self._save('sensor_heatmap_bottom')

### 2018-07-10 doc/plos/code/ugly_plot.py
flip heatmap top and bottom

@@ -22,7 +22,7 @@ class UglyPlot:
     ax.set_xticklabels(['%d' % weeks[i] for i in wk_idx])
     ax.set_yticklabels(sensors[::-1])
     plt.pcolor(data[::-1, :])
-    self._save('sensor_heatmap_top')
+    self._save('sensor_heatmap_bottom')
 
     for i, sensor in enumerate(sensors):
       plt.plot(data[i, :], label=sensor)
@@ -30,4 +30,4 @@ class UglyPlot:
     plt.plot([t[w] for w in weeks], label='Ground Truth', color='black')
     plt.ylim([0, 10])
     plt.legend()
-    self._save('sensor_heatmap_bottom')
+    self._save('sensor_heatmap_top')

### 2018-07-10 doc/plos/code/analyze.py
drop nowcasts without sensors

@@ -60,6 +60,18 @@ class Analysis:
     result = {}
     for week, place, value, std in self.data['nc_%s' % name]:
       if place == loc:
+        if (loc == 'vi' and week < 201327) or (loc == 'pr' and week < 201453):
+          # The nowcast for this location on this week was made *without* any
+          # sensors for this location on this week. It's only available by
+          # inference indirectly from sensors in other locations in the same
+          # region. That these nowcasts were possible at all is a testament to
+          # the flexibility of sensor fusion, but they are not interesting from
+          # the perspective of comparing nowcast accuracy to sensor accuracy --
+          # there are no sensors with which to compare. Full disclosure, these
+          # nowcasts are much less accurate than nowcasts for which direct
+          # sensors are available, and this is reflected by their associated
+          # standard deviations on these weeks.
+          continue
         result[week] = value
     return result
 

### 2018-07-10 doc/plos/text/manuscript.tex
add figure 1 to draft
make preview figures larger
figure 2: all nowcasts

@@
 
+% Don't include figures in the final version. This is just for preview.
+\newgeometry{top=1.0in,left=1.0in,footskip=30pt}
+\begin{figure}[p]
+  \includegraphics[width=\textwidth]{../data/figures/sensor_heatmap_top.png}
+  Figure 1 (top).
+\end{figure}
+\begin{figure}[p]
+  \includegraphics[width=\textwidth]{../data/figures/sensor_heatmap_bottom.png}
+  Figure 1 (bottom).
+\end{figure}
+\begin{figure}[p]
+  \includegraphics[width=\textwidth]{../data/figures/all_nowcasts.png}
+  Figure 2.
+\end{figure}
+
 \end{document}

### 2018-07-10 doc/plos/code/analyze.py

@@ -113,6 +113,7 @@ def main():
   analysis = Analysis()
   plotter = UglyPlot(analysis)
   plotter.plot_sensor_heatmap()
+  plotter.plot_all_nowcasts()
 
 
 if __name__ == '__main__':

### 2018-07-10 doc/plos/code/ugly_plot.py

@@ -3,6 +3,9 @@
 import numpy as np
 import matplotlib.pylab as plt
 
+from delphi.utils import epiweek as Epiweek
+from delphi.utils.geo.locations import Locations
+
 
 class UglyPlot:
 
@@ -15,6 +18,7 @@ class UglyPlot:
 
   def plot_sensor_heatmap(self):
     data, sensors, weeks = self.analysis.get_heatmap_data()
+
     fig, ax = plt.subplots()
     wk_idx = list(range(0, len(weeks), 100))
     plt.xticks(wk_idx)
@@ -31,3 +35,29 @@ class UglyPlot:
     plt.ylim([0, 10])
     plt.legend()
     self._save('sensor_heatmap_top')
+
+  def plot_all_nowcasts(self):
+
+    def get_xy(data):
+      weeks = sorted(data)
+      y = [data[w] for w in weeks]
+      x = [Epiweek.delta_epiweeks(201030, w) for w in weeks]
+      return list(map(np.array, [x, y]))
+
+    def shift(i, x, y):
+      x = x / max(x)
+      y = y / max(y)
+      x = x * 0.8 + 0.1
+      y = y * 0.8 + 0.1
+      r = i // 5
+      c = i % 5
+      return x + c, y - r, r, c
+
+    for i, loc in enumerate(Locations.region_list):
+      x, y, r, c = shift(i, *get_xy(self.analysis.get_truth(loc)))
+      s, t, r, c = shift(i, *get_xy(self.analysis.get_nowcast(loc)))
+      x, y = x[-len(s):], y[-len(t):]
+      plt.plot(x, y, color='#404040')
+      plt.plot(s, t, color='#008060')
+      plt.text(c, -r + 0.8, loc)
+    self._save('all_nowcasts')

### 2018-07-11 doc/plos/code/analyze.py
add plot accuracy_vs_sensors
also tweak the size of all the other plots

@@ -114,6 +114,7 @@ def main():
   plotter = UglyPlot(analysis)
   plotter.plot_sensor_heatmap()
   plotter.plot_all_nowcasts()
+  plotter.plot_accuracy_vs_sensors()
 
 
 if __name__ == '__main__':

### 2018-07-11 doc/plos/code/ugly_plot.py

@@ -3,6 +3,7 @@
 import numpy as np
 import matplotlib.pylab as plt
 
+from delphi.nowcast.util.flu_data_source import FluDataSource
 from delphi.utils import epiweek as Epiweek
 from delphi.utils.geo.locations import Locations
 
@@ -19,7 +20,7 @@ class UglyPlot:
   def plot_sensor_heatmap(self):
     data, sensors, weeks = self.analysis.get_heatmap_data()
 
-    fig, ax = plt.subplots()
+    fig, ax = plt.subplots(figsize=(6, 3))
     wk_idx = list(range(0, len(weeks), 100))
     plt.xticks(wk_idx)
     plt.yticks([0.5 + i for i in range(len(sensors))])
@@ -28,6 +29,7 @@ class UglyPlot:
     plt.pcolor(data[::-1, :])
     self._save('sensor_heatmap_bottom')
 
+    plt.figure(figsize=(6, 3))
     for i, sensor in enumerate(sensors):
       plt.plot(data[i, :], label=sensor)
     t = self.analysis.get_truth('nat')
@@ -53,6 +55,7 @@ class UglyPlot:
       c = i % 5
       return x + c, y - r, r, c
 
+    plt.figure(figsize=(6, 8))
     for i, loc in enumerate(Locations.region_list):
       x, y, r, c = shift(i, *get_xy(self.analysis.get_truth(loc)))
       s, t, r, c = shift(i, *get_xy(self.analysis.get_nowcast(loc)))
@@ -61,3 +64,33 @@ class UglyPlot:
       plt.plot(s, t, color='#008060')
       plt.text(c, -r + 0.8, loc)
     self._save('all_nowcasts')
+
+  def plot_accuracy_vs_sensors(self):
+    fig, axes = plt.subplots(nrows=4, ncols=2, figsize=(6, 8))
+
+    loc = 'nat'
+    for i, sensor in enumerate(FluDataSource.SENSORS):
+      t = self.analysis.get_truth(loc)
+      n = self.analysis.get_nowcast(loc)
+      s = self.analysis.get_sensor(sensor, loc)
+      ns = self.analysis.merge(n, s)
+      t = dict((w, t[w]) for w in ns.keys())
+      n = dict((w, n[w]) for w in ns.keys())
+      s = dict((w, s[w]) for w in ns.keys())
+      ts = self.analysis.merge(t, s)
[…]

### 2018-07-11 doc/plos/code/analyze.py
add plot accuracy_vs_ablation

@@ -115,6 +115,7 @@ def main():
   plotter.plot_sensor_heatmap()
   plotter.plot_all_nowcasts()
   plotter.plot_accuracy_vs_sensors()
+  plotter.plot_accuracy_vs_ablation()
 
 
 if __name__ == '__main__':

### 2018-07-11 doc/plos/code/ugly_plot.py

@@ -94,3 +94,33 @@ class UglyPlot:
       axes[r, c].text(.4, 3.5, '%s (%d)' % (sensor, st_n['n']))
 
     self._save('accuracy_vs_sensors')
+
+  def plot_accuracy_vs_ablation(self):
+    fig, axes = plt.subplots(nrows=4, ncols=2, figsize=(6, 8))
+
+    loc = 'nat'
+    for i, sensor in enumerate(FluDataSource.SENSORS):
+      t = self.analysis.get_truth(loc)
+      n = self.analysis.get_nowcast(loc)
+      s = self.analysis.get_experiment('ablate_%s' % sensor, loc)
+      ns = self.analysis.merge(n, s)
+      t = dict((w, t[w]) for w in ns.keys())
+      n = dict((w, n[w]) for w in ns.keys())
+      s = dict((w, s[w]) for w in ns.keys())
+      ts = self.analysis.merge(t, s)
+      tn = self.analysis.merge(t, n)
+
+      st_s = self.analysis.get_stats(ts)
+      st_n = self.analysis.get_stats(tn)
+      if st_s['n'] != st_n['n']:
+        raise Exception()
+
+      r = i // 2
+      c = i % 2
+      x = [3, 4, 0, 1]
+      y = [st_s['mae'], st_n['mae'], st_s['rmse'], st_n['rmse']]
+      axes[r, c].barh(x, y, tick_label=['-', '+', '-', '+'])
+      axes[r, c].set_xlim([0, 0.3])
+      axes[r, c].text(.2, 3.5, '%s (%d)' % (sensor, st_n['n']))
+
+    self._save('accuracy_vs_ablation')

### 2018-07-11 doc/plos/code/analyze.py
add plot accuracy_vs_abscission

@@ -116,6 +116,7 @@ def main():
   plotter.plot_all_nowcasts()
   plotter.plot_accuracy_vs_sensors()
   plotter.plot_accuracy_vs_ablation()
+  plotter.plot_accuracy_vs_abscission()
 
 
 if __name__ == '__main__':

### 2018-07-11 doc/plos/code/ugly_plot.py

@@ -124,3 +124,37 @@ class UglyPlot:
       axes[r, c].text(.2, 3.5, '%s (%d)' % (sensor, st_n['n']))
 
     self._save('accuracy_vs_ablation')
+
+  def plot_accuracy_vs_abscission(self):
+    fig, ax = plt.subplots(figsize=(3, 3))
+
+    loc = 'nat'
+    tru = self.analysis.get_truth(loc)
+    a_n = self.analysis.get_experiment('abscise2_national', loc)
+    a_r = self.analysis.get_experiment('abscise2_regional', loc)
+    a_s = self.analysis.get_experiment('abscise2_state', loc)
+
+    tvn = self.analysis.merge(tru, a_n)
+    tvr = self.analysis.merge(tru, a_r)
+    tvs = self.analysis.merge(tru, a_s)
+
+    st_n = self.analysis.get_stats(tvn)
+    st_r = self.analysis.get_stats(tvr)
+    st_s = self.analysis.get_stats(tvs)
+    if st_n['n'] != st_r['n'] or st_n['n'] != st_s['n']:
+      raise Exception()
+
+    x = [6, 5, 4, 2, 1, 0]
+    y = [
+      st_n['mae'],
+      st_r['mae'],
+      st_s['mae'],
+      st_n['rmse'],
+      st_r['rmse'],
+      st_s['rmse'],
+    ]
+    ax.barh(x, y, tick_label=['Nm', 'Rm', 'Sm', 'Nr', 'Rr', 'Sr'])
+    ax.set_xlim([0, 0.3])
+    ax.text(.25, 5, '(%d)' % (st_n['n']))
+
+    self._save('accuracy_vs_abscission')

### 2018-07-12 doc/plos/text/manuscript.tex
add plot accuracy_vs_sensors
also tweak the size of all the other plots
add plot accuracy_vs_ablation
add plot accuracy_vs_abscission
covairance wording and references
I tried to make the description of the covariance estimation method
clearer, and I added references to published methods.

@@
 
-Estimating covariance in this situation is challenging, and while the
-statistical literature suggests several approaches TODO(cite a few of them),
-consensus appears to be lacking.
+Estimating covariance in this situation is challenging, and the statistical
+literature suggests several approaches \cite{bickel2008covariance,
+ollerer2015robust, bien2011sparse, friedman2008sparse, bickel2008regularized,
+hsieh2011sparse}. For speed and simplicity, we follow the guidance of
+\cite{kolar2012consistent} for handling missing values and of
+\cite{touloumis2015nonparametric} for regularization.
 
-We begin by computing the empirical pair-wise covariance of all sensors.
-However, the resulting matrix does not necessarily describe multi-dimensional
-covariance (i.e. it may not be positive definite). To correct this, we shrink
-this matrix toward the diagonal matrix of sensor variance. The magnitude of the
-shrinkage weight is chosen to maximize the multivariate normal log-likelihood
-of the available data. We subsequently use the shrunken matrix as $R$ in sensor
-fusion.
+We begin by building a matrix of pair-wise empirical covariance. In this
+matrix, the value of each element is computed as the covariance between a pair
+of sensors. By filling the matrix, covariance is computed independently for all
+sensor pairs. However, the resulting matrix does not necessarily describe
+multi-dimensional covariance (i.e. it may not be positive definite), especially
+with many missing values. To handle this, we shrink this matrix toward the
+diagonal matrix of sensor variance. The magnitude of the shrinkage weight is
+chosen to maximize the multivariate normal log-likelihood of the available
+data. We subsequently use the shrunken matrix as $R$ in sensor fusion.
 
@@
 
+\bibitem{bickel2008covariance}
+Bickel PJ, Levina E.
+\newblock Covariance regularization by thresholding.
+\newblock The Annals of Statistics. 2008; p. 2577--2604.
+
+\bibitem{ollerer2015robust}
+{\"O}llerer V, Croux C.
+\newblock Robust high-dimensional precision matrix estimation.
+\newblock In: Modern Nonparametric, Robust and Multivariate Methods. Springer;
+  2015. p. 325--350.
+
+\bibitem{bien2011sparse}
+Bien J, Tibshirani RJ.
+\newblock Sparse estimation of a covariance matrix.
+\newblock Biometrika. 2011;98(4):807--820.
+
+\bibitem{friedman2008sparse}
+Friedman J, Hastie T, Tibshirani R.
+\newblock Sparse inverse covariance estimation with the graphical lasso.
+\newblock Biostatistics. 2008;9(3):432--441.
+
+\bibitem{bickel2008regularized}
+Bickel PJ, Levina E.
+\newblock Regularized estimation of large covariance matrices.
+\newblock The Annals of Statistics. 2008; p. 199--227.
+
+\bibitem{hsieh2011sparse}
+Hsieh CJ, Dhillon IS, Ravikumar PK, Sustik MA.
+\newblock Sparse inverse covariance matrix estimation using quadratic
+  approximation.
+\newblock In: Advances in Neural Information Processing Systems; 2011. p.
+  2330--2338.
+
+\bibitem{kolar2012consistent}
+Kolar M, Xing EP.
+\newblock Consistent covariance selection from data with missing values.
+\newblock In: Proceedings of the 29th International Conference on Machine
+  Learning (ICML-12); 2012. p. 551--558.
+
+\bibitem{touloumis2015nonparametric}
+Touloumis A.
+\newblock Nonparametric Stein-type shrinkage covariance matrix estimators in
+  high-dimensional settings.
+\newblock Computational Statistics \& Data Analysis. 2015;83:251--261.
+
 \end{thebibliography}
@@
 \end{figure}
+\begin{figure}[p]
+  \includegraphics[width=\textwidth]{../data/figures/accuracy_vs_sensors.png}
+  Figure 3.
+\end{figure}
+\begin{figure}[p]
+  \includegraphics[width=\textwidth]{../data/figures/accuracy_vs_ablation.png}
+  Figure 4.
+\end{figure}
+\begin{figure}[p]
+  \includegraphics[width=\textwidth]{../data/figures/accuracy_vs_abscission.png}
+  Figure 5.
+\end{figure}
 

### 2018-07-12 doc/plos/code/analyze.py
basic sensor and nowcast stats

@@ -1,3 +1,4 @@
+import argparse
 import csv
 import glob
 import os
@@ -110,13 +111,50 @@ class Analysis:
 
 
 def main():
+  parser = argparse.ArgumentParser()
+  parser.add_argument(
+      'name',
+      help='name of the analysis to run')
+  args = parser.parse_args()
+  name = args.name
+
   analysis = Analysis()
-  plotter = UglyPlot(analysis)
-  plotter.plot_sensor_heatmap()
-  plotter.plot_all_nowcasts()
-  plotter.plot_accuracy_vs_sensors()
-  plotter.plot_accuracy_vs_ablation()
-  plotter.plot_accuracy_vs_abscission()
+
+  if name == 'nowcast_info':
+    n = analysis.get_nowcast('nat')
+    print('num national nowcasts: %d' % len(n))
+    print('first week: %d' % min(n))
+    print('last week: %d' % max(n))
+    t = 0
+    for l in Locations.region_list:
+      t += len(analysis.get_nowcast(l))
+    print('total num nowcasts: %d' % t)
+    print('num locations: %d' % len(Locations.region_list))
+  elif name == 'sensor_info':
+    grand_total = 0
+    for s in FluDataSource.SENSORS:
+      print('%s:' % s)
+      n = analysis.get_sensor(s, 'nat')
+      print('  num national: %d' % len(n))
+      print('  first week: %d' % min(n))
+      print('  last week: %d' % max(n))
+      num_loc = 0
+      for l in Locations.region_list:
+        n = len(analysis.get_sensor(s, l))
+        if n:
+          num_loc += 1
+        grand_total += n
+      print('  num locations: %d' % num_loc)
+    print('total num readings: %d' % grand_total)
+  elif name == 'plot':
+    plotter = UglyPlot(analysis)
+    plotter.plot_sensor_heatmap()
+    plotter.plot_all_nowcasts()
+    plotter.plot_accuracy_vs_sensors()
+    plotter.plot_accuracy_vs_ablation()
+    plotter.plot_accuracy_vs_abscission()
+  else:
+    raise Exception('unknown analysis: %s' % name)
 
 
 if __name__ == '__main__':

### 2018-07-12 doc/plos/text/manuscript.tex
start results section
Add some basic stats, a table, and make figures in-line with text.
Figures will be removed from the final version.

@@
 \newcommand{\ipsum}{{\bf IPSUM}}
+
+% TODO: don't include graphics in the final version
+\newcommand{\maybeincludegraphics}[1]{\includegraphics[width=\textwidth]{../data/figures/#1.png}}
 
@@
 
-TODO
+\subsection*{Sensors}
+
+% python3 analyze.py sensor_info
+% gft:
+%   num national: 254
+%   first week: 201040
+%   last week: 201532
+%   num locations: 62
+% ght:
+%   num national: 402
+%   first week: 201040
+%   last week: 201824
+%   num locations: 52
+% twtr:
+%   num national: 339
+%   first week: 201151
+%   last week: 201824
+%   num locations: 71
+% wiki:
+%   num national: 402
+%   first week: 201040
+%   last week: 201824
+%   num locations: 1
+% cdc:
+%   num national: 276
+%   first week: 201310
+%   last week: 201824
+%   num locations: 71
+% epic:
+%   num national: 119
+%   first week: 201442
+%   last week: 201819
+%   num locations: 15
+% sar3:
+%   num national: 402
+%   first week: 201040
+%   last week: 201824
+%   num locations: 75
+% arch:
+%   num national: 265
+%   first week: 201040
+%   last week: 201820
+%   num locations: 20
+% total num readings: 113164
+
+In total, we obtained 113,164 individual sensor readings from the eight
+surveillance sources discussed above. Table~\ref{table_sensor_breakdown} gives
+a breakdown of the number of weeks and locations covered by each source.
+Fig~\ref{figure_sensor_heatmap} shows time-series of sensor readings for the US
+as a whole, which is the only location common to all data sources.
+
+\begin{table}[!ht]
+  \centering
+  \caption{{\bf TODO(Table caption.)}}
+  \begin{tabular}{|l|l|l|}
+    \hline
+    {\bf Source} & {\bf Weeks} & {\bf Locations}\\ \thickhline
+    GFT & 254 & 62 \\ \hline
+    GHT & 402 & 52 \\ \hline
+    TWTR & 339 & 71 \\ \hline
+    WIKI & 402 & 1 \\ \hline
+    CDCP & 276 & 71 \\ \hline
+    EPIC & 119 & 15 \\ \hline
+    SAR3 & 402 & 75 \\ \hline
+    ARCH & 265 & 20 \\ \hline
+  \end{tabular}
+  \begin{flushleft}
+    TODO(Table notes.)
+  \end{flushleft}
+  \label{table_sensor_breakdown}
+\end{table}
+
+\begin{figure}[!h]
+  \maybeincludegraphics{sensor_heatmap_top}
+  \maybeincludegraphics{sensor_heatmap_bottom}
+  \caption{
+    {\bf TODO(figure title.)}
+    TODO(Figure caption top and bottom.)
+  }
+  \label{figure_sensor_heatmap}
+\end{figure}
+
+\subsection*{Nowcasts}
+
+% python3 analyze.py nowcast_info
+% num national nowcasts: 397
+% first week: 201045
+% last week: 201824
+% total num nowcasts: 29103
+% num locations: 75
+
+In total, we produced 29,103 nowcasts for 75 US locations across 397 weeks
+spanning 2010w45--2018w24. Locations include the US as a whole, the 10 HHS
+regions, the 9 census divisions, the 50 states, the District of Columbia,
+Puerto Rico, the US Virgin Islands, New York City, and New York state
+\textit{excluding} New York City. Although these were computed in retrospect,
+models were trained using only data that would have been available at the time.
+Namely, whenever possible, preliminary values of wILI were used in place of
+final values.
 
@@
 
-% Don't include figures in the final version. This is just for preview.
-\newgeometry{top=1.0in,left=1.0in,footskip=30pt}
-\begin{figure}[p]
-  \includegraphics[width=\textwidth]{../data/figures/sensor_heatmap_top.png}
-  Figure 1 (top).
-\end{figure}
-\begin{figure}[p]
-  \includegraphics[width=\textwidth]{../data/figures/sensor_heatmap_bottom.png}
-  Figure 1 (bottom).
-\end{figure}
-\begin{figure}[p]
-  \includegraphics[width=\textwidth]{../data/figures/all_nowcasts.png}
-  Figure 2.
-\end{figure}
-\begin{figure}[p]
-  \includegraphics[width=\textwidth]{../data/figures/accuracy_vs_sensors.png}
-  Figure 3.
-\end{figure}
-\begin{figure}[p]
-  \includegraphics[width=\textwidth]{../data/figures/accuracy_vs_ablation.png}
-  Figure 4.
-\end{figure}
-\begin{figure}[p]
-  \includegraphics[width=\textwidth]{../data/figures/accuracy_vs_abscission.png}
-  Figure 5.
-\end{figure}
-
 \end{document}

### 2018-07-13 doc/plos/text/manuscript.tex
combine sensor plots
The top/bottom subplots are now a single image. I updated the title and
caption of Table 1 and Figure 1.

@@
   \centering
-  \caption{{\bf TODO(Table caption.)}}
+  \caption{{\bf Sensor coverage across time and location.}}
   \begin{tabular}{|l|l|l|}
@@
   \end{tabular}
-  \begin{flushleft}
-    TODO(Table notes.)
-  \end{flushleft}
   \label{table_sensor_breakdown}
@@
 \begin{figure}[!h]
-  \maybeincludegraphics{sensor_heatmap_top}
-  \maybeincludegraphics{sensor_heatmap_bottom}
+  \maybeincludegraphics{sensor_heatmap}
   \caption{
-    {\bf TODO(figure title.)}
-    TODO(Figure caption top and bottom.)
+    {\bf National sensor readings fitted from eight data sources.}
+    Top: Plot of sensor readings and ground truth wILI (black) over time,
+    highlighting sensor noise.
+    Bottom: Heatmap of sensor readings over time, highlighting sensor
+    intermittency.
   }

### 2018-07-13 doc/plos/code/ugly_plot.py

@@ -20,23 +20,25 @@ class UglyPlot:
   def plot_sensor_heatmap(self):
     data, sensors, weeks = self.analysis.get_heatmap_data()
 
-    fig, ax = plt.subplots(figsize=(6, 3))
+    fig, axes = plt.subplots(nrows=2, ncols=1, figsize=(6, 6))
+
+    ax = axes[1]
     wk_idx = list(range(0, len(weeks), 100))
-    plt.xticks(wk_idx)
-    plt.yticks([0.5 + i for i in range(len(sensors))])
+    ax.set_xticks(wk_idx)
+    ax.set_yticks([0.5 + i for i in range(len(sensors))])
     ax.set_xticklabels(['%d' % weeks[i] for i in wk_idx])
     ax.set_yticklabels(sensors[::-1])
-    plt.pcolor(data[::-1, :])
-    self._save('sensor_heatmap_bottom')
+    ax.pcolor(data[::-1, :])
 
-    plt.figure(figsize=(6, 3))
+    ax = axes[0]
     for i, sensor in enumerate(sensors):
-      plt.plot(data[i, :], label=sensor)
+      ax.plot(data[i, :], label=sensor)
     t = self.analysis.get_truth('nat')
-    plt.plot([t[w] for w in weeks], label='Ground Truth', color='black')
-    plt.ylim([0, 10])
-    plt.legend()
-    self._save('sensor_heatmap_top')
+    ax.plot([t[w] for w in weeks], label='Ground Truth', color='black')
+    ax.set_ylim([0, 10])
+    ax.legend()
+
+    self._save('sensor_heatmap')
 
   def plot_all_nowcasts(self):
 

### 2018-07-13 doc/plos/text/manuscript.tex
add all_nowcasts figure and caption

@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
   \maybeincludegraphics{sensor_heatmap}
@@
     {\bf National sensor readings fitted from eight data sources.}
-    Top: Plot of sensor readings and ground truth wILI (black) over time,
-    highlighting sensor noise.
+    Top: Plot of sensor readings (colors) and ground truth wILI (black) over
+    time, highlighting sensor noise.
     Bottom: Heatmap of sensor readings over time, highlighting sensor
@@
 Puerto Rico, the US Virgin Islands, New York City, and New York state
-\textit{excluding} New York City. Although these were computed in retrospect,
-models were trained using only data that would have been available at the time.
-Namely, whenever possible, preliminary values of wILI were used in place of
-final values.
+\textit{excluding} New York City. Fig~\ref{figure_all_nowcasts} shows a
+time-series comparison of all nowcasts with ground truth. Although these
+nowcasts were computed in retrospect, models were trained using only data that
+would have been available at the time. Namely, whenever possible, preliminary
+values of wILI were used in place of final values.
+
+\begin{figure}[!ht]
+  \maybeincludegraphics{all_nowcasts}
+  \caption{
+    {\bf Nowcasts compared with ground truth.}
+    Each subpanel contains a plot of nowcast (color) and ground truth (black)
+    in a particular location within the US. All subplots share the same scale
+    in both time and magnitude.
+  }
+  \label{figure_all_nowcasts}
+\end{figure}
 

### 2018-07-13 doc/plos/text/manuscript.tex
add subsection evaluation

@@
 
+\subsection*{Evaluation}
+
+In a given location over a given period of time, we assess the accuracy of a
+set of nowcasts by two measures: mean absolute error (MAE) and root mean
+squared error (RMSE). With ground truth vector $\vec{x}$, nowcast vector
+$\vec{y}$, and elements indexed by week $w$ of $n$, these measures are defined
+as:
+
+\begin{align*}
+  \text{MAE} &= \frac{1}{n} \sum_{w=1}^n{\left|x_w - y_w\right|} \\
+  \text{RMSE} &= \left( \frac{1}{n} \sum_{w=1}^n{\left(x_w - y_w\right)^2} \right)^\frac{1}{2}
+\end{align*}
+
 % Results and Discussion can be combined.

### 2018-07-17 doc/plos/text/manuscript.tex
add accuracy measure MASE
See also:
Hyndman RJ, Koehler AB. Another look at measures of forecast accuracy.
International journal of forecasting. 2006 Oct 1;22(4):679-88.

@@
 
-In a given location over a given period of time, we assess the accuracy of a
-set of nowcasts by two measures: mean absolute error (MAE) and root mean
+In a given location over a given period of time, we assess the standalone
+accuracy of a nowcast by two measures: mean absolute error (MAE) root mean
 squared error (RMSE). With ground truth vector $\vec{x}$, nowcast vector
@@
   \text{RMSE} &= \left( \frac{1}{n} \sum_{w=1}^n{\left(x_w - y_w\right)^2} \right)^\frac{1}{2}
+\end{align*}
+
+Although the units of ground truth (i.e. wILI) are the same in all locations,
+the normal range of wILI varies between locations. To facilitate comparison
+between locations, we also report mean absolute scaled error (MASE)
+\cite{hyndman2006another}. MASE is the MAE of a given nowcast normalized by the
+MAE of a naive nowcast in the same location. We define the naive nowcast for
+any given location on any given week as the reported ground truth in the same
+location 52 weeks earlier. Using the same notation as above, and considering
+the $m$ weeks for which ground truth is available for at least 52 weeks prior,
+MASE is defined as:
+
+\begin{align*}
+  \text{MASE} &= \text{MAE} \left( \frac{1}{m} \sum_{w=1}^m{\left|x_w - x_{w - 52}\right|} \right) ^ {-1}
 \end{align*}
@@
 
+\bibitem{hyndman2006another}
+Hyndman RJ, Koehler AB.
+\newblock Another look at measures of forecast accuracy.
+\newblock International journal of forecasting. 2006;22(4):679--688.
+
 \end{thebibliography}

### 2018-07-17 doc/plos/code/analyze.py

@@ -83,12 +83,24 @@ class Analysis:
     return dict((ew, (a[ew], b[ew])) for ew in a.keys() & b.keys())
 
   def get_stats(self, merged):
-    a = np.array([merged[ew][0] for ew in merged])
-    b = np.array([merged[ew][1] for ew in merged])
+    a = np.array([merged[ew][0] for ew in sorted(merged)])
+    b = np.array([merged[ew][1] for ew in sorted(merged)])
+    mae = np.mean(np.abs(a - b))
+
+    # normalize MAE by MAE of naive nowcaster
+    naive = []
+    for ew1 in merged:
+      ew0 = Epiweek.add_epiweeks(ew1, -52)
+      if ew0 in merged:
+        naive.append(np.abs(merged[ew1][0] - merged[ew0][0]))
+    mae_naive = np.mean(naive)
+
     return {
       'n': len(merged),
       'rmse': np.sqrt(np.mean(np.square(a - b))),
-      'mae': np.mean(np.abs(a - b)),
+      'mae': mae,
+      'mae_naive': mae_naive,
+      'mase': mae / mae_naive,
     }
 
   def get_heatmap_data(self):

### 2018-07-23 doc/plos/code/analyze.py
add plot all_mase

@@ -165,6 +165,7 @@ def main():
     plotter.plot_accuracy_vs_sensors()
     plotter.plot_accuracy_vs_ablation()
     plotter.plot_accuracy_vs_abscission()
+    plotter.plot_all_mase()
   else:
     raise Exception('unknown analysis: %s' % name)
 

### 2018-07-23 doc/plos/code/ugly_plot.py

@@ -16,6 +16,7 @@ class UglyPlot:
   def _save(self, name):
     plt.savefig('data/figures/' + name + '.png', bbox_inches='tight')
     plt.close()
+    print('>', name)
 
   def plot_sensor_heatmap(self):
     data, sensors, weeks = self.analysis.get_heatmap_data()
@@ -160,3 +161,32 @@ class UglyPlot:
     ax.text(.25, 5, '(%d)' % (st_n['n']))
 
     self._save('accuracy_vs_abscission')
+
+  def plot_all_mase(self):
+    fig, ax = plt.subplots(figsize=(6, 3))
+    nums = []
+    colors = []
+    for i, loc in enumerate(Locations.region_list):
+      idx = i % 2
+      idx += 2 if loc in Locations.hhs_list else 0
+      idx += 4 if loc in Locations.cen_list else 0
+      idx += 6 if loc in Locations.atom_list else 0
+      colors.append([
+        '#404040',
+        '#808080',
+        '#804040',
+        '#e08080',
+        '#408040',
+        '#80e080',
+        '#404080',
+        '#8080e0'][idx])
+      ft = self.analysis.get_truth(loc)
+      nc = self.analysis.get_nowcast(loc)
+      merged = self.analysis.merge(ft, nc)
+      st = self.analysis.get_stats(merged)
+      nums.append(st['mase'])
+    idx = list(range(len(nums)))
+    plt.bar(idx, nums, color=colors)
+    ax.tick_params(axis='x', labelsize=5)
+    plt.xticks(idx, Locations.region_list, rotation='vertical')
+    self._save('all_mase')

### 2018-07-24 doc/plos/code/analyze.py
comparison table, rmsse, naive, alternative
- Includes naive and alternative nowcasts.
- The `get_stats` method has been replaced by the `get_metrics` method.
- Added  metric RMSSE (from the same paper as MASE).
- Generate comparison table for all nowcasters and most measures.

@@ -79,29 +79,45 @@ class Analysis:
   def get_nowcast(self, loc):
     return self.get_experiment('vanilla', loc)
 
-  def merge(self, a, b):
-    return dict((ew, (a[ew], b[ew])) for ew in a.keys() & b.keys())
-
-  def get_stats(self, merged):
-    a = np.array([merged[ew][0] for ew in sorted(merged)])
-    b = np.array([merged[ew][1] for ew in sorted(merged)])
-    mae = np.mean(np.abs(a - b))
-
-    # normalize MAE by MAE of naive nowcaster
-    naive = []
-    for ew1 in merged:
+  def get_naive_nowcast(self, loc):
+    nowcast = {}
+    truth = self.get_truth(loc)
+    for ew1 in truth:
       ew0 = Epiweek.add_epiweeks(ew1, -52)
-      if ew0 in merged:
-        naive.append(np.abs(merged[ew1][0] - merged[ew0][0]))
-    mae_naive = np.mean(naive)
+      if ew0 in truth:
+        nowcast[ew1] = truth[ew0]
+    return nowcast
 
-    return {
-      'n': len(merged),
+  def get_alternative_nowcast(self, loc):
+    sensors = {}
+    for sensor in FluDataSource.SENSORS:
+      sensors[sensor] = self.get_sensor(sensor, loc)
+    nowcast = {}
+    for ew in self.get_truth(loc):
+      values = []
+      for sensor in FluDataSource.SENSORS:
+        if ew in sensors[sensor]:
+          values.append(sensors[sensor][ew])
+      if values:
+        nowcast[ew] = np.median(values)
+    return nowcast
+
+  def get_metrics(self, truth, nowcast, naive=None):
+    common_weeks = sorted(truth.keys() & nowcast.keys())
+    a = np.array([truth[ew] for ew in common_weeks])
+    b = np.array([nowcast[ew] for ew in common_weeks])
+    result = {
+      'n': len(common_weeks),
       'rmse': np.sqrt(np.mean(np.square(a - b))),
-      'mae': mae,
-      'mae_naive': mae_naive,
-      'mase': mae / mae_naive,
+      'mae': np.mean(np.abs(a - b)),
     }
+    if naive is not None:
+      common_weeks = sorted(nowcast.keys() & naive.keys())
+      trimmed_naive = dict([(ew, naive[ew]) for ew in common_weeks])
+      result_naive = self.get_metrics(truth, trimmed_naive)
[…]

### 2018-07-24 doc/plos/code/ugly_plot.py

@@ -76,15 +76,12 @@ class UglyPlot:
       t = self.analysis.get_truth(loc)
       n = self.analysis.get_nowcast(loc)
       s = self.analysis.get_sensor(sensor, loc)
-      ns = self.analysis.merge(n, s)
-      t = dict((w, t[w]) for w in ns.keys())
-      n = dict((w, n[w]) for w in ns.keys())
-      s = dict((w, s[w]) for w in ns.keys())
-      ts = self.analysis.merge(t, s)
-      tn = self.analysis.merge(t, n)
+      common = n.keys() & s.keys()
+      n = dict((w, n[w]) for w in common)
+      s = dict((w, s[w]) for w in common)
 
-      st_s = self.analysis.get_stats(ts)
-      st_n = self.analysis.get_stats(tn)
+      st_s = self.analysis.get_metrics(t, s)
+      st_n = self.analysis.get_metrics(t, n)
       if st_s['n'] != st_n['n']:
         raise Exception()
 
@@ -106,15 +103,12 @@ class UglyPlot:
       t = self.analysis.get_truth(loc)
       n = self.analysis.get_nowcast(loc)
       s = self.analysis.get_experiment('ablate_%s' % sensor, loc)
-      ns = self.analysis.merge(n, s)
-      t = dict((w, t[w]) for w in ns.keys())
-      n = dict((w, n[w]) for w in ns.keys())
-      s = dict((w, s[w]) for w in ns.keys())
-      ts = self.analysis.merge(t, s)
-      tn = self.analysis.merge(t, n)
+      common = n.keys() & s.keys()
+      n = dict((w, n[w]) for w in common)
+      s = dict((w, s[w]) for w in common)
 
-      st_s = self.analysis.get_stats(ts)
-      st_n = self.analysis.get_stats(tn)
+      st_s = self.analysis.get_metrics(t, s)
+      st_n = self.analysis.get_metrics(t, n)
       if st_s['n'] != st_n['n']:
         raise Exception()
 
@@ -137,13 +131,9 @@ class UglyPlot:
     a_r = self.analysis.get_experiment('abscise2_regional', loc)
     a_s = self.analysis.get_experiment('abscise2_state', loc)
 
-    tvn = self.analysis.merge(tru, a_n)
-    tvr = self.analysis.merge(tru, a_r)
-    tvs = self.analysis.merge(tru, a_s)
-
-    st_n = self.analysis.get_stats(tvn)
[…]

### 2018-07-24 doc/plos/code/analyze.py
introduce concept of naive oracle
And update analysis to use this new definition of a "naive" nowcast.

@@ -80,15 +80,38 @@ class Analysis:
     return self.get_experiment('vanilla', loc)
 
   def get_naive_nowcast(self, loc):
+    # Because final wILI is not known for multiple months, it's not possible to
+    # implement a *real-time* random walk naive nowcaster. There are (at least)
+    # two ways to define a substitute naive nowcaster:
+    #   - Naive Oracle: assume final wILI is known at runtime (it's not)
+    #     and define the nowcast as final wILI on the previous week.
+    #   - Seasonal Naive: define the nowcast as final wILI on the same week
+    #     one year in the past.
+    # Naive Oracle has the disadvantage that it's not realistic (because of
+    # backfill), and therefore it is unfairly advantaged. Seasonal Naive has
+    # the disadvantage that wILI 52 weeks ago is only very loosely correlated
+    # with wILI at runtime, and therefore it is unfairly disadvantaged.
+    # (Ideally we would define the naive nowcast as preliminary wILI on the
+    # previous week, but that data isn't generally available, except for
+    # certain locations and seasons.)
+    # It's not immediately clear which definition of "naive" is better in this
+    # situation. The variable below controls which definition is used
+    # throughout this analysis; 1 corresponds to Naive Oracle, and 52
+    # corresponds to Seasonal Naive.
+    delta = 1
+
     nowcast = {}
     truth = self.get_truth(loc)
     for ew1 in truth:
-      ew0 = Epiweek.add_epiweeks(ew1, -52)
+      ew0 = Epiweek.add_epiweeks(ew1, -delta)
       if ew0 in truth:
         nowcast[ew1] = truth[ew0]
     return nowcast
 
   def get_alternative_nowcast(self, loc):
+    # For each week, report the median value of all sensor readings in this
+    # location, ignoring sensor readings in all other locations. This is a
+    # basic "wisdom of the crowd" approach to nowcasting.
     sensors = {}
     for sensor in FluDataSource.SENSORS:
       sensors[sensor] = self.get_sensor(sensor, loc)

### 2018-07-24 doc/plos/text/manuscript.tex
discuss Naive Oracle in manuscript

@@
 \cite{hyndman2006another}. MASE is the MAE of a given nowcast normalized by the
-MAE of a naive nowcast in the same location. We define the naive nowcast for
-any given location on any given week as the reported ground truth in the same
-location 52 weeks earlier. Using the same notation as above, and considering
-the $m$ weeks for which ground truth is available for at least 52 weeks prior,
-MASE is defined as:
+MAE of a one-step naive nowcast computed in-sample in the same location.
+
+Because wILI is subject to revision over multiple months, a real-time naive
+nowcast would be defined as \textit{preliminary} wILI on the preceding week.
+However, preliminary versions of wILI are not generally known today (except for
+within a narrow set of locations and seasons). As a result, we introduce the
+``Naive Oracle'' nowcaster. This assumes (incorrectly) that the final version
+of wILI is known with no delay and defines the naive nowcast as \textit{final}
+wILI on the preceding week. Since this information is not available at runtime,
+the Naive Oracle is unfairly advantaged. As such, MASE is reasonably subject to
+being greater than one, indicating that the nowcast under consideration
+performed \textit{worse} than the ``naive'' nowcast. This should be
+unsurprising, if not expected, given that the Naive Oracle incorporates
+information which is not available at runtime. In any case, MASE is a useful
+measure for comparing the performance of a given nowcast across different
+locations.
+
+Given the above discussion of the Naive Oracle, we define the naive nowcast for
+any given location on any given week as the final value of ground truth in the
+same location one week earlier. Using the same notation as above, and
+considering the $m$ weeks for which ground truth is available on the prior
+week, MASE is defined as:
 
 \begin{align*}
-  \text{MASE} &= \text{MAE} \left( \frac{1}{m} \sum_{w=1}^m{\left|x_w - x_{w - 52}\right|} \right) ^ {-1}
+  \text{MASE} &= \text{MAE} \left( \frac{1}{m} \sum_{w=1}^m{\left|x_w - x_{w - 1}\right|} \right) ^ {-1}
 \end{align*}

### 2018-07-25 doc/plos/text/manuscript.tex
MASE rewording

@@
 unsurprising, if not expected, given that the Naive Oracle incorporates
-information which is not available at runtime. In any case, MASE is a useful
-measure for comparing the performance of a given nowcast across different
-locations.
+information which is not available at runtime. In any case, while MASE is not
+particularly helpful as an absolute measure of accuracy, MASE is useful for
+comparing relative accuracy between different locations.
 
-Given the above discussion of the Naive Oracle, we define the naive nowcast for
-any given location on any given week as the final value of ground truth in the
-same location one week earlier. Using the same notation as above, and
-considering the $m$ weeks for which ground truth is available on the prior
-week, MASE is defined as:
+We define the naive nowcast for any given location on any given week as the
+final value of ground truth in the same location one week earlier. Using the
+same notation as above, and considering the $m$ weeks for which ground truth is
+available on the prior week, MASE is defined as:
 

### 2018-07-25 doc/plos/code/analyze.py
export error table as CSV
Includes for each location:
- MAE Actual
- MAE Baseline
- RMSE Actual
- RMSE Baseline
- MASE Actual

@@ -1,5 +1,6 @@
 import argparse
 import csv
+import functools
 import glob
 import os
 
@@ -160,6 +161,25 @@ class Analysis:
 
     return data, FluDataSource.SENSORS, weeks
 
+  @functools.lru_cache()
+  def get_accuracy_table(self):
+    result = {}
+    for loc in Locations.region_list:
+      t = self.get_truth(loc)
+      n_actual = self.get_nowcast(loc)
+      n_alt = self.get_alternative_nowcast(loc)
+      n_naive = self.get_naive_nowcast(loc)
+      st_actual = self.get_metrics(t, n_actual, naive=n_naive)
+      st_alt = self.get_metrics(t, n_alt, naive=n_naive)
+      result[loc] = {
+        'actual_mae': st_actual['mae'],
+        'actual_rmse': st_actual['rmse'],
+        'actual_mase': st_actual['mase'],
+        'alt_mae': st_alt['mae'],
+        'alt_rmse': st_alt['rmse'],
+      }
+    return result
+
 
 def main():
   parser = argparse.ArgumentParser()
@@ -209,29 +229,29 @@ def main():
     plotter.plot_all_mase()
 
   if name in ('all', 'table'):
-    fmt = '%s' + ' & %.3f' * 10 + ' \\\\ \\hline'
-    for loc in Locations.region_list:
-      t = analysis.get_truth(loc)
-      n_actual = analysis.get_nowcast(loc)
-      n_alt = analysis.get_alternative_nowcast(loc)
-      n_naive = analysis.get_naive_nowcast(loc)
-      st_actual = analysis.get_metrics(t, n_actual, naive=n_naive)
-      st_alt = analysis.get_metrics(t, n_alt, naive=n_naive)
-      st_naive = analysis.get_metrics(t, n_naive, naive=n_naive)
-      args = [
-        loc,
-        st_actual['mae'],
-        st_alt['mae'],
-        st_naive['mae'],
-        st_actual['mase'],
-        st_alt['mase'],
-        st_actual['rmse'],
-        st_alt['rmse'],
-        st_naive['rmse'],
-        st_actual['rmsse'],
-        st_alt['rmsse'],
-      ]
-      print(fmt % tuple(args))
+    filename = 'data/files/table.csv'
+    print('saving %s' % filename)
+    with open(filename, 'w') as f:
+      writer = csv.writer(f)
+      writer.writerow([
+        'Location',
+        'MAE Actual',
[…]

### 2018-07-25 doc/plos/text/manuscript.tex
add RelMAE, reword MASE

--- 2018-07-24 doc/plos/text/manuscript.tex
@@
 
-In a given location over a given period of time, we assess the standalone
-accuracy of a nowcast by two measures: mean absolute error (MAE) root mean
-squared error (RMSE). With ground truth vector $\vec{x}$, nowcast vector
-$\vec{y}$, and elements indexed by week $w$ of $n$, these measures are defined
-as:
+In a given location over a given period of time, we assess the absolute
+(scale-dependent) error of a nowcast by two measures: mean absolute error (MAE)
+and root mean squared error (RMSE). With ground truth vector $\vec{x}$, nowcast
+vector $\vec{y}$, and elements indexed by week $w$ of $n$, these measures are
+defined as:
 
@@
 Although the units of ground truth (i.e. wILI) are the same in all locations,
-the normal range of wILI varies between locations. To facilitate comparison
-between locations, we also report mean absolute scaled error (MASE)
-\cite{hyndman2006another}. MASE is the MAE of a given nowcast normalized by the
-MAE of a one-step naive nowcast computed in-sample in the same location.
+the normal range of wILI varies between locations. To facilitate aggregation
+and comparison among locations, we report several relative (scale-independent)
+error measures with respect to various baselines. Each of these is computed by
+normalizing nowcast error by the same type of error measured on the output of a
+baseline model. Here we define two such baseline models: ``naive'' and
+``simple''.
 
@@
 However, preliminary versions of wILI are not generally known today (except for
-within a narrow set of locations and seasons). As a result, we introduce the
-``Naive Oracle'' nowcaster. This assumes (incorrectly) that the final version
+within a narrow set of locations and seasons). As a result, we introduce a
+``naive oracle'' nowcaster. This assumes (incorrectly) that the final version
 of wILI is known with no delay and defines the naive nowcast as \textit{final}
 wILI on the preceding week. Since this information is not available at runtime,
-the Naive Oracle is unfairly advantaged. As such, MASE is reasonably subject to
-being greater than one, indicating that the nowcast under consideration
-performed \textit{worse} than the ``naive'' nowcast. This should be
-unsurprising, if not expected, given that the Naive Oracle incorporates
-information which is not available at runtime. In any case, MASE is a useful
-measure for comparing the performance of a given nowcast across different
+the naive baseline is unfairly advantaged.
+
+The ``simple'' baseline is defined as the median of all sensor readings for a
+given location, ignoring sensor readings elsewhere. This is in effect a wisdom
+of the crowd nowcast that doesn't incorporate information from outside
 locations.
 
-Given the above discussion of the Naive Oracle, we define the naive nowcast for
-any given location on any given week as the final value of ground truth in the
-same location one week earlier. Using the same notation as above, and
-considering the $m$ weeks for which ground truth is available on the prior
-week, MASE is defined as:
+We report MAE of nowcasts relative to MAE of each baseline (RelMAE TODO(cite)).
+In doing so with the ``naive'' baseline, this is a special case of RelMAE known
+as mean absolute scaled error (MASE) \cite{hyndman2006another}. MASE is defined
+in the context of time-series forecasting in general as the MAE of a given forecast
+normalized by the MAE of an in-sample, one-step naive forecast.
+
+Using the same notation as above, assuming $\tilde{z}_w$ represents the median
+sensor reading for the current location on week $w$, and considering the $m$
+weeks for which ground truth is available on the prior week, these measures are
+defined as:
 
 \begin{align*}
+  \text{RelMAE} &= \text{MAE} \left( \frac{1}{n} \sum_{w=1}^n{\left|x_w - \tilde{z}_w\right|} \right) ^ {-1} \\
   \text{MASE} &= \text{MAE} \left( \frac{1}{m} \sum_{w=1}^m{\left|x_w - x_{w - 1}\right|} \right) ^ {-1}
@@
 
+MASE is reasonably subject to being greater than one, indicating that the
+nowcast under consideration performed \textit{worse} than the ``naive''
+nowcast. This should be unsurprising, if not expected, given that the naive
+oracle incorporates information which is not available at runtime.
+
 TODO

### 2018-08-01 doc/plos/code/analyze.py
analysis: update table, add metrics
This shows geometric mean of MASE/RelMAE/RelMASE "horizontally" (nat,
regions, states) and "vertically" (states within each region).

@@ -171,12 +171,17 @@ class Analysis:
       n_naive = self.get_naive_nowcast(loc)
       st_actual = self.get_metrics(t, n_actual, naive=n_naive)
       st_alt = self.get_metrics(t, n_alt, naive=n_naive)
+      st_naive = self.get_metrics(t, n_naive, naive=n_naive)
       result[loc] = {
         'actual_mae': st_actual['mae'],
         'actual_rmse': st_actual['rmse'],
         'actual_mase': st_actual['mase'],
+        'actual_relmae': st_actual['mae'] / st_alt['mae'],
         'alt_mae': st_alt['mae'],
         'alt_rmse': st_alt['rmse'],
+        'alt_mase': st_alt['mase'],
+        'naive_mae': st_naive['mae'],
+        'naive_rmse': st_naive['rmse'],
       }
     return result
 
@@ -237,9 +242,12 @@ def main():
         'Location',
         'MAE Actual',
         'MAE Baseline',
+        'MAE Naive',
         'RMSE Actual',
         'RMSE Baseline',
+        'RMSE Naive',
         'MASE Actual',
+        'MASE Baseline',
       ])
       table = analysis.get_accuracy_table()
       for loc in Locations.region_list:
@@ -247,11 +255,77 @@ def main():
           loc,
           '%.3f' % table[loc]['actual_mae'],
           '%.3f' % table[loc]['alt_mae'],
+          '%.3f' % table[loc]['naive_mae'],
           '%.3f' % table[loc]['actual_rmse'],
           '%.3f' % table[loc]['alt_rmse'],
+          '%.3f' % table[loc]['naive_rmse'],
           '%.3f' % table[loc]['actual_mase'],
+          '%.3f' % table[loc]['alt_mase'],
         ])
     print('saved %s' % filename)
+
+  if name in ('all', 'metrics'):
+    table = analysis.get_accuracy_table()
+    rel = {'mae': {}, 'rmse': {}, 'mase': {}}
+    for loc in Locations.region_list:
+      rel['mae'][loc] = table[loc]['actual_mae'] / table[loc]['alt_mae']
+      rel['rmse'][loc] = table[loc]['actual_rmse'] / table[loc]['alt_rmse']
+      rel['mase'][loc] = table[loc]['actual_mase']
+
+    def geo_mean(locs, metric):
[…]

### 2018-08-01 doc/plos/text/manuscript.tex
add accuracy_vs_sensors figure
Some slight rewording, add new subsubsections.

@@
 as mean absolute scaled error (MASE) \cite{hyndman2006another}. MASE is defined
-in the context of time-series forecasting in general as the MAE of a given forecast
-normalized by the MAE of an in-sample, one-step naive forecast.
+in the context of time-series forecasting in general as the MAE of a given
+forecast normalized by the MAE of an in-sample, one-step naive forecast.
 
@@
 time-series comparison of all nowcasts with ground truth. Although these
-nowcasts were computed in retrospect, models were trained using only data that
-would have been available at the time. Namely, whenever possible, preliminary
-values of wILI were used in place of final values.
+nowcasts were computed in retrospect, models were trained whenever possible
+using only data that would have been available at the time. For any given week,
+preliminary values of wILI (when known) were used in place of final values, and
+data from future weeks was held out.
 
@@
 
+\subsection*{Accuracy relative to data sources}
+
+The error (measured in MAE and RMSE) of nowcasts of ILI for the US as a whole
+relative to the same error for each data source are shown in
+Fig~\ref{figure_accuracy_vs_sensors}.
+
+\begin{figure}[!ht]
+  \maybeincludegraphics{accuracy_vs_sensors}
+  \caption{
+    {\bf Nowcast error compared with sensor error.}
+    TODO: explain this
+  }
+  \label{figure_accuracy_vs_sensors}
+\end{figure}
+
+\subsection*{Accuracy relative to baselines}
+
+TODO (python3 analyze.py metrics)
+
 \section*{Discussion}

### 2018-08-01 doc/plos/text/manuscript.tex
commentary on accuracy vs sensors

@@
 
-\subsection*{Accuracy relative to data sources}
-
-The error (measured in MAE and RMSE) of nowcasts of ILI for the US as a whole
-relative to the same error for each data source are shown in
-Fig~\ref{figure_accuracy_vs_sensors}.
+\subsubsection*{Accuracy relative to data sources}
+
+For the US as a whole, Fig~\ref{figure_accuracy_vs_sensors} compares error
+(measured in MAE and RMSE) between nowcasts and individual data sources. To
+control for the fact that difficulty varies over time, the evaluation is
+performed only over weeks on which the nowcast and the particular data source
+are jointly available. The number of weeks in each such union is indicated in
+each corresponding subpanel.
 
@@
 
-\subsection*{Accuracy relative to baselines}
+In all instances, the nowcast has lower error (both MAE and RMSE) than any
+individual data source, subject to the fitting procedure described above. This
+indicates, at least for the US as a whole, that the nowcast is on average more
+accurate than any individual data source.
+
+\subsubsection*{Accuracy relative to baselines}
 
@@
 oracle incorporates information which is not available at runtime.
+
+That the nowcast is more accurate than its individual inputs should hold in
+general as the outputs of sensor fusion should, on average, have smaller error
+than the inputs. However, this may not hold in all cases due to overfitting or
+violations of the assumptions of linearity and normality.
 

### 2018-08-01 doc/plos/text/manuscript.tex
wording on error relative to baseline

@@
 
+To assess accuracy at finer geographic resolutions, we report MAE relative to
+the naive and simple baselines (measured as the ratios MASE and RelMAE
+respectively). When aggregating errors over multiple geographic units (e.g.
+states within some region) we report the geometric mean of errors as this
+yields a result that can still be interpreted as a ratio.
+
 TODO (python3 analyze.py metrics)

### 2018-08-07 doc/plos/code/analyze.py
generate relative metrics table

@@ -327,6 +327,19 @@ def main():
       val = geo_mean_str(Locations.cen_map[reg], rel['mase'])
       print('atoms in %s gm MASE: %s' % (reg, val))
 
+    def print_row(name, locs):
+      n = len(locs)
+      mase = geo_mean(locs, rel['mase'])
+      relmae = geo_mean(locs, rel['mae'])
+      print('%s (%d) & %.3f & %.3f \\\\ \\hline' % (name, n, mase, relmae))
+    print('-' * 20)
+    print_row('US National', Locations.nat_list)
+    print_row('Census Divisions', Locations.cen_list)
+    print_row('HHS Regions', Locations.hhs_list)
+    for reg in Locations.hhs_list:
+      print_row('States in %s' % reg.upper(), Locations.hhs_map[reg])
+    print_row('All States', Locations.atom_list)
+
 
 if __name__ == '__main__':
   main()

### 2018-08-07 doc/plos/text/manuscript.tex
add relative metrics table
And some commentary.

@@
 states within some region) we report the geometric mean of errors as this
-yields a result that can still be interpreted as a ratio.
-
-TODO (python3 analyze.py metrics)
+yields a result that can still be interpreted as a ratio. The relative errors
+are summarized in Table~\ref{table_relative_metrics}.
+
+% python3 analyze.py metrics
+\begin{table}[!ht]
+  \centering
+  \caption{{\bf Relative error across sets of locations.}}
+  \begin{tabular}{|l|l|l|}
+    \hline
+    {\bf Locations (n)} & {\bf G.M. MASE} & {\bf G.M. RelMAE}\\ \thickhline
+    US National (1) & 0.667 & 0.799 \\ \hline
+    Census Divisions (9) & 0.853 & 0.859 \\ \hline
+    HHS Regions (10) & 0.865 & 0.849 \\ \hline
+    Atoms in HHS1 (6) & 0.960 & 0.918 \\ \hline
+    Atoms in HHS2 (5) & 1.094 & 1.068 \\ \hline
+    Atoms in HHS3 (6) & 1.019 & 0.900 \\ \hline
+    Atoms in HHS4 (8) & 0.996 & 0.886 \\ \hline
+    Atoms in HHS5 (6) & 0.955 & 0.912 \\ \hline
+    Atoms in HHS6 (5) & 0.941 & 0.941 \\ \hline
+    Atoms in HHS7 (4) & 0.995 & 0.928 \\ \hline
+    Atoms in HHS8 (6) & 1.029 & 0.991 \\ \hline
+    Atoms in HHS9 (4) & 1.048 & 0.863 \\ \hline
+    Atoms in HHS10 (4) & 1.014 & 0.929 \\ \hline
+    All Atoms (54) & 1.002 & 0.931 \\ \hline
+  \end{tabular}
+  \begin{flushleft}
+    TODO(define atom, enumerate "all atoms", mention naive/simple, define GM)
+  \end{flushleft}
+  \label{table_relative_metrics}
+\end{table}
+
+At the national level, the sensor fusion nowcast has lower error than either
+baseline by a considerable margin. The same is true for Census divisions and
+HHS regions collectively, although the margin is not as large. Aggregated over
+individual atoms, either within a particular region or altogether, the sensor
+fusion nowcast seems to generally be on par with the naive baseline and
+somewhat better than the simple baseline. For atoms within HHS region 2
+specifically, the sensor fusion nowcast has higher overall error than both
+naive and simple baselines.
 
@@
 
-TODO
+TODO(why is hhs2 the worst? PR/VI: least ground truth, extreme noise, disparate
+dynamics, no sensors other than sar3; NY, JFK: inferred.)
+
+TODO(more discussion)
 

### 2018-08-13 doc/plos/text/manuscript.tex
add ablation experiments

@@
 
+\subsubsection*{Sensitivity to each data source}
+
+Digital surveillance is intermittent by nature, and data sources are subject to
+sporadic, periodic (e.g. EPIC), and permanent (e.g. GFT) periods of
+absence. We expect sensor fusion to gracefully handle these cases, but it is
+not immediately clear what impact the absence of a given data source has on the
+quality of the resulting nowcast.
+
+To explore this question, we compare accuracy of nowcasts with and without each
+data source in a set of eight ``ablation'' experiments in
+Fig~\ref{figure_accuracy_vs_ablation}. For a given data source, we compute the
+US national nowcast with and without that source and compare error in terms of
+MAE and RMSE. Since the difficulty of nowcasting varies over time, we restrict
+these comparisons to the set of weeks on which a particular data source (or,
+more accurately, its fitted sensors) are available.
+
+\begin{figure}[!ht]
+  \maybeincludegraphics{accuracy_vs_ablation}
+  \caption{
+    {\bf Nowcast error compared with and without each data source.}
+    TODO: explain this
+  }
+  \label{figure_accuracy_vs_ablation}
+\end{figure}
+
+As expected, we generally find that removal of a particular data source causes
+an increase (or, in the case of WIKI, no meaningful change) in nowcasting
+error. However, there are a couple of exceptions: TWTR and CDC. Nowcasting
+error \textit{decreases} when removing these data sources. This is unexpected
+in theory and strongly suggests violated assumptions (e.g. the relationship of
+these sources with wILI may be nonlinear) or overfitting (e.g. covariance may
+be underestimated for these sources).
+
+\subsubsection*{Sensitivity to geographic resolution}
+
+TODO(abscission experiments)
+
 \section*{Discussion}
@@
 dynamics, no sensors other than sar3; NY, JFK: inferred.)
+
+TODO(curse of dimensionality and regularization)
+
+TODO(ablation: overfitting, assumptions, time-variable relationships)
 

### 2018-09-04 doc/plos/text/manuscript.tex
add abscission experiments
and some small wording tweaks

@@
 In all instances, the nowcast has lower error (both MAE and RMSE) than any
-individual data source, subject to the fitting procedure described above. This
+individual data source after applying the fitting procedure described above. This
 indicates, at least for the US as a whole, that the nowcast is on average more
-accurate than any individual data source.
+accurate than any individual data source (at least for this particular fitting
+procedure).
 
@@
 in theory and strongly suggests violated assumptions (e.g. the relationship of
-these sources with wILI may be nonlinear) or overfitting (e.g. covariance may
-be underestimated for these sources).
+these sources with wILI may be nonlinear or residuals may not be normally
+distributed) or overfitting (e.g. covariance may be underestimated for these
+sources) in practice.
 
@@
 
-TODO(abscission experiments)
+A related question is to what extent the inclusion of geographic ``tiers''
+impacts accuracy of nowcasts. Namely, is it helpful for nowcasting the US
+nationally to include data at the level of regions, states, cities, and
+territories?
+
+To answer this, we similarly compare nowcasting error in terms of MAE and RMSE
+with and without data at regional (Census and HHS) and state (including cities
+and territories) levels. This is difficult to assess for all data sources,
+given that many do not include data at the level of states (or even regions, in
+the case of WIKI). Because of this, we limit the set of data sources to only
+those available at the level of states (and above): GFT, GHT, TWTR, CDCP, and
+SAR3. Again, we restrict the comparison to weeks on which sensors are mutually
+available. The results of these comparisons are summarized in
+Fig~\ref{figure_accuracy_vs_abscission}.
+
+\begin{figure}[!ht]
+  \maybeincludegraphics{accuracy_vs_abscission}
+  \caption{
+    {\bf Nowcast error compared across geographic tiers.}
+    TODO: explain this
+  }
+  \label{figure_accuracy_vs_abscission}
+\end{figure}
+
+Between national-only and national-and-regional nowcasts, we see an increase in
+accuracy (a decrease in both MAE and RMSE), suggesting that higher geographic
+resolution is beneficial to the nowcast. When including all locations, MAE
+further decreases, but RMSE increases. One possible reason for this mixed
+result is poor model fitting (i.e. covariance estimation) induced by a large
+increase in dimensionality with no increase in the number of weekly sensor
+observations.
 
@@
 \label{S1_Text}
-{\bf Supporting Information.} Includes derivation and sensitivity analysis.
+{\bf Supporting Information.} Derivation of the sensor fusion kernel.
 

### 2018-09-05 doc/plos/text/manuscript.tex
discuss things

@@
 
-MASE is reasonably subject to being greater than one, indicating that the
-nowcast under consideration performed \textit{worse} than the ``naive''
-nowcast. This should be unsurprising, if not expected, given that the naive
-oracle incorporates information which is not available at runtime.
-
-That the nowcast is more accurate than its individual inputs should hold in
-general as the outputs of sensor fusion should, on average, have smaller error
-than the inputs. However, this may not hold in all cases due to overfitting or
-violations of the assumptions of linearity and normality.
+We compared nowcasts with two baselines; a naive oracle, which reports
+finalized wILI on the previous week, and a simple method, which reports the
+median sensor reading on a given week. At high geographic resolution (i.e.
+states, territories, and cities), we find that sensor fusion nowcasts are
+cumulatively approximately on par with naive oracle nowcasts and somewhat
+better than simple nowcasts. At lower geographic resolution (i.e. regionally
+and nationally), sensor fusion nowcasts strongly and uniformly outperform both
+baselines.
+
+MASE---error of sensor fusion nowcasts relative to error of the naive
+oracle---is frequently greater than one in small geographic units, indicating
+that the nowcast under consideration performed \textit{worse} than the
+so-called ``naive'' nowcast. This result is not entirely unexpected, given that
+the naive oracle incorporates latent information (namely, finalized ILI) which
+is unavailable to operational nowcasts at runtime.
+
+That the nowcast is more accurate than its individual inputs (i.e. ``sensors'')
+should hold in general as the outputs of sensor fusion should, on average, have
+smaller error than the inputs. Overfitting or violations of the assumptions of
+linearity and normality may cause deviations from this expectation.
+
+We observe such deviations at least twice. First, for the set of five states,
+territories, and cities within HHS region two (``HHS2''), we find that the
+geometric mean of RelMAE---error of sensor fusion nowcasts relative to error of
+the simple baseline---is greater than one, meaning that the baseline
+outperformed the nowcast in these locations. Second, we find that removal of
+certain sensors (``ablation'' and ``abscission'' experiments) causes the
+accuracy of the nowcast to improve. In both of these situations, less data is
+better than more data, against expectation.
+
+Our estimate of the covariance of sensor ``noise'' (residuals w.r.t. ground
+truth) $R$ is necessarily imperfect. Adding additional sensors exponentially
+increases the dimension of the problem without adding new observations---the
+curse of dimensionality. With 367 sensor-location pairs, $R$ has shape $367
+\times 376$; yet in the best case we only have a few hundred observations from
+which to estimate this covariance---often much fewer. Our only recourse is
+aggressive regularization at the expense of covariance accuracy, and therefore
+nowcast accuracy.
+
+Linearity between sensors and ground truth and normality of sensor noise are
+each merely assumed. Violations of these assumptions are neither tested for nor
+acted upon. The sensor fitting procedure simply centers and scales raw data to
+fit ground truth in a least-squares sense. It is possible, if not likely, that
+some sensors are nonlinear w.r.t. ground truth or have non-normal residuals (or
+both). When this is the case, nowcasts will be less accurate; sensor fusion, at
+least as formulated above, operates in terms of linear combinations of
+multivariate normal distributions.
+
+Additionally, we make the implicit assumption that the relationship between
+each sensor and ground truth is constant over time. That is, we assume that the
+bias and variance of residuals are fixed. (In contrast, we explicitly account
+for a time-varying relationship between raw data and ground truth in the sensor
+fitting procedure.) However, this is almost certainly violated in some
+situations; for example, digital surveillance is disproportionately affected
+during periods of heightened traditional and social media coverage of
+outbreaks.
 
@@
 
-TODO(curse of dimensionality and regularization)
-
-TODO(ablation: overfitting, assumptions, time-variable relationships)
-
-TODO(more discussion)
+TODO: merits and drawbacks of sensor fusion
 

### 2018-09-12 doc/plos/text/manuscript.tex
correction for number of sensors

@@
 increases the dimension of the problem without adding new observations---the
-curse of dimensionality. With 367 sensor-location pairs, $R$ has shape $367
-\times 376$; yet in the best case we only have a few hundred observations from
-which to estimate this covariance---often much fewer. Our only recourse is
+curse of dimensionality. With up to 363 sensor-location pairs, $R$ has shape
+$363 \times 363$; yet in the best case we only have a few hundred observations
+from which to estimate this covariance---often much fewer. Our only recourse is
 aggressive regularization at the expense of covariance accuracy, and therefore

### 2018-10-09 doc/plos/text/manuscript.tex
clarify dimensionality

--- 2018-09-05 doc/plos/text/manuscript.tex
@@
 Our estimate of the covariance of sensor ``noise'' (residuals w.r.t. ground
-truth) $R$ is necessarily imperfect. Adding additional sensors exponentially
+truth) $R$ is necessarily imperfect. Adding additional sensors quadratically
 increases the dimension of the problem without adding new observations---the
-curse of dimensionality. With 367 sensor-location pairs, $R$ has shape $367
-\times 376$; yet in the best case we only have a few hundred observations from
-which to estimate this covariance---often much fewer. Our only recourse is
-aggressive regularization at the expense of covariance accuracy, and therefore
-nowcast accuracy.
+curse of dimensionality. With up to 363 sensor-location pairs, $R$ has shape
+$363 \times 363$; yet in the best case we only have a few hundred weekly
+observations from which to estimate this covariance---often much fewer. Our
+only recourse is aggressive regularization at the expense of covariance
+accuracy, and therefore nowcast accuracy.
 

### 2018-10-10 doc/plos/text/manuscript.tex
discuss relative inaccuracy in hhs2

@@
 
+There are several challenges specific to nowcasting in HHS2 and Census region
+two. First, the constituent territories Puerto Rico (PR) and the US Virgin
+Islands (VI) have limited ground truth---far fewer observations than all other
+locations. Second, the only sensor currently available for each of New York
+\textit{excluding New York City}, PR, and VI is SAR3---a time-series model
+which incorporates no real-world measurements. Finally, the epidemiological
+dynamics (timing, strain distribution, etc.) within HHS2 are disparate; New
+Jersey and New York (including New York City) form one epidemiological cluster
+while PI and VI form another.
+
 Our estimate of the covariance of sensor ``noise'' (residuals w.r.t. ground
@@
 
-TODO(why is hhs2 the worst? PR/VI: least ground truth, extreme noise, disparate
-dynamics, no sensors other than sar3; NY, JFK: inferred.)
-
 TODO: merits and drawbacks of sensor fusion

### 2018-10-10 doc/plos/text/manuscript.tex
discuss pros and cons of sensor fusion

@@
 
-TODO: merits and drawbacks of sensor fusion
+Herein we used a data assimilation technique which we refer to as ``sensor
+fusion''. As a consequence of being derived from the Kalman filter, sensor
+fusion has a number of advantages, including minimizing the expectation of
+squared error, robustness to partial and missing observations, and a natural
+framework for assimilating data across overlapping geographic locations.
+However, sensor fusion ultimately defines a nowcast as a simple linear
+combination of the available sensors. When assumptions of linearity and
+normality are violated, and when covariance estimates are inaccurate, the
+accuracy of resulting nowcasts suffers.
 

### 2018-10-10 doc/plos/text/manuscript.tex
concluding remarks

@@
 
-TODO
+Here we took an approach to nowcasting flu in quasi-real-time using a method
+based on the Kalman filter. We used this method to assimilate digital
+surveillance and flu forecasts from eight sources spanning 75 distinct
+locations within the US.
+
+We showed that accuracy of nowcasts almost always outperforms straightforward
+baselines. The advantage of these nowcasts is their extremely low
+latency---they are available up to one week before initial wILI publication.
+
+The primary challenge in incorporating additional data sources is accurately
+estimating sensor noise covariance. The number of weekly observations required
+to estimate covariance at the same resolution grows with the square of the
+number of input sensors. To address this problem, a direction for future work
+is that of sensor selection---perhaps via LASSO.
+
+Another area of future research is to nowcast additional outbreaks in
+additional locations. This, however, depends critically on the availability of
+both a rich history of ground truth and an ongoing set of digital surveillance
+sources.
+
+Finally, in addition to retrospective nowcasts examined here, ongoing nowcasts
+are produced each week and made publicly available both through our website and
+our developer API (TODO: citations).
 
